---
title: "Reproduce TE"
author: "Christine Lucille Kuryla"
date: "2024-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)

# import data

te_df_original <- read_rds("./TE_df_original.RDS")

# for 49 sample size
hr_raw <- read_csv("./data/data_from_email/hr_cleaned.csv")
bp_raw <- read_csv("./data/data_from_email/bp_cleaned.csv")
rr_raw <- read_csv("./data/data_from_email/rsp_cleaned.csv")
pt_cond_raw <- read_csv("./data/data_from_email/patient condition.csv")

# For 108 sample size
# hr_raw <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/raw_data_110_202404/hr_raw_108.csv")
# rr_raw <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/raw_data_110_202404/rr_raw_105.csv")
# bp_raw <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/raw_data_110_202404/bp_raw_107.csv")

# for graphing
# string vector
periods_str <- c("Base",
                 "Pre",
                 "Instruct",
                 "PrepTask",
                 "Task",
                 "Post5",
                 "Post10",
                 "Post20",
                 "Post30",
                 "Post60",
                 "Post90",
                 "Post120",
                 "DBT",
                 "ST",
                 "SST",
                 "CP")

# factor vector
# create a vector with the periods as factors in the correct order (not alphabetical)
# this is for ease of plotting later
periods_factor <- factor(periods_str, levels = periods_str)

```

Hard code missing data

```{r missing_data}

# See import RMD for details
# For now, if there exists at least one missing modality, we will remove that id/period combo.

missing_data <- (rbind(
  c("mi002", "PrepTask"), #rsp 
  c("mi002", "Task"), #rsp
c("mi009", "Base"), #rttimes
c("mi012", "Instruct"), #bptimes
c("mi012", "Post5"), #bptimes
c("mi012", "Post120"), #bptimes
c("mi012", "DBT"), #bptimes
c("mi012", "ST"), #bptimes
c("mi016", "Instruct"), # rttimes for Instruct missing
c("mi024", "Post120"), # bptimes for Post120 missing
c("mi028", "SST"), # rttimes for SST missing
c("mi032", "Post10"), # rsp for Post10 missing
c("mi041", "Post120"), # mi041Post120_bptimes.csv 
c("mi046", "Post5"), # mi046Post5_rttimes.csv 
c("mi050", "ST"), # mi050ST_bptimes.csv 
c("mi053", "Post60"), # mi053Post60_rsp.csv 
c("mi083", "SST"), # mi083SST_rttimes.csv , mi083SST_rsp.csv 
c("mi087", "DBT"), # mi087DBT_bptimes.csv 
c("mi088", "ST"), # mi088ST_bptimes.csv 
c("mi098", "Post30"), # mi098Post30_bptimes.csv 
c("mi105", "Task"), # mi105Task_rsp.csv 
c("mi108", "PrepTask"), # mi108PrepTask_rsp.csv 
c("mi110", "Post5"), # mi110Post5_rsp.csv , mi110Post10_rsp.csv , mi110Post120_bptimes.csv 
c("mi110", "Post10"),
c("mi110", "Post120"),
c("mi109", "CP"), # mi109 is missing rsp for CP only
c("mi070", "CP"), # mi070 is missing rttimes for CP only
c("mi026", "CP"), # mi026 is missing bptimes for CP only but there is another file mi026CP.2.bi --> check what this means
c("mi094", "CP"), # mi094 is missing the last 4 tasks (CP, SST, ST, DBT)
c("mi094", "SST"),
c("mi094", "ST"),
c("mi094", "DBT"),
c("mi104", "CP"), # mi104 is missing the last 4 tasks (CP, SST, ST, DBT)
c("mi104", "SST"),
c("mi104", "ST"),
c("mi104", "DBT"),
c("mi106", "CP"), # mi106 is missing all CP, rttimes in Task, and bptimes in DBT
c("mi106", "Task"),
c("mi106", "DBT"),
c("mi069", "Pre") # missing BP
))

colnames(missing_data) <- c("id", "period")

# also just take out mi003, mi004, mi047

# Summary

# mi003 is missing rsp
# mi004 is missing rsp
# mi047 is missing rsp and bptimes

# mi014 does not have correct files (it just says acquisition?) --> omitted
# mi075 is missing completely


```

Initial EDA 

```{r initial_eda}

# converting names from the subsample 
hr_raw <- hr_raw %>% 
  mutate(heart.rate = rate, heart.time = time) %>% 
  filter(!(period %in% c("C3", "Acquisition", "NA")))
rr_raw <- rr_raw %>% 
  mutate(rsp.rate = rate, rsp.time = time) %>% 
  filter(!(period %in% c("C3", "Acquisition", "NA")))
bp_raw <- bp_raw %>% 
  mutate(bp.pVal = pVal, bp.time = pSec) %>% 
  filter(!(period %in% c("C3", "Acquisition", "NA")))

# Visualize the raw data

ggplot(hr_raw, aes(x = heart.time, y = heart.rate, group = id, color = condition)) + 
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Heart Rate")

ggplot(rr_raw, aes(x = rsp.time, y = rsp.rate, group = id, color = condition)) + 
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Respiration Rate")

ggplot(bp_raw, aes(x = bp.time, y = bp.pVal, group = id, color = condition)) +
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + 
  ggtitle("Blood Pressure")
```


# Map data to matching intervals and combine

```{r map_to_matching_intervals}
# Assuming you have the following data frames:
# - hr_df: heart rate data with columns 'id', 'period', 'time', 'rate'
# - rr_df: respiration data with columns 'id', 'period', 'pSec', 'pVal', 'rate'
# - bp_df: blood pressure data with columns 'id', 'period', 'pSec', 'pVal'

# hr_df <- hr_raw %>%
#   mutate(time = heart.time) %>%
#   mutate(rate = heart.rate) %>%
#   select(id, period, condition, time, rate)
# 
# rr_df <- rr_raw %>% 
#   mutate(pSec = rsp.pSec) %>% 
#   mutate(pVal = rsp.pVal) %>% 
#   mutate(rate = rsp.rate) %>% 
#   select(id, period, condition, pSec, pVal, rate)
# 
# bp_df <- bp_raw %>% 
#   mutate(pSec = bptimes.pSec) %>% 
#   mutate(pVal = bp.pVal) %>% 
#   select(id, period, condition, pSec, pVal)

# hr_df <- hr_raw
# rr_df <- rsp_raw #rr_raw
# bp_df <- bp_raw

hr_df <- hr_raw 
rr_df <- rr_raw %>% 
  mutate(rsp.pSec = rsp.time,
         rsp.pVal = pVal) 
bp_df <- bp_raw %>% 
  mutate(bptimes.pSec = bp.time) 
  

# Desired time interval (in seconds)
interval <- 1

# # Get unique subject IDs and periods
# 
# subject_ids <- unique(c(hr_df$id, rr_df$id, bp_df$id))
# periods <- unique(c(hr_df$period, rr_df$period, bp_df$period))


subject_ids <- intersect(hr_df$id, rr_df$id)
subject_ids<- intersect(subject_ids, bp_df$id)
periods <- intersect(hr_df$period, rr_df$period)
periods <- intersect(periods, bp_df$period)

# This is for reproducibility test, don't actually do this
subject_ids <- unique(te_df_original %>% pull(id))

library(future)
future::plan(multisession)

# Initialize an empty list to store the aligned data frames
#aligned_data <- list()
aligned_data <- data.frame()

# Iterate over each subject and period
for (subject in subject_ids) {
  for (period in periods) {
    # Subset data for the current subject and period
    hr_sub <- hr_df %>% filter(id == subject, period == period) 
    rsp_sub <- rr_df %>% filter(id == subject, period == period) 
    bp_sub <- bp_df %>% filter(id == subject, period == period)
    
    # Find the common time range
    min_start_time <- min(hr_sub$heart.time, rsp_sub$rsp.pSec, bp_sub$bptimes.pSec, na.rm = TRUE)
   #max_end_time <- max(hr_sub$heart.time, rsp_sub$rsp.pSec, bp_sub$bptimes.pSec, na.rm = TRUE)
    min_end_time <- min(max(hr_sub$heart.time, na.rm = TRUE), 
                        max(rsp_sub$rsp.pSec, na.rm = TRUE), 
                        max(bp_sub$bptimes.pSec, na.rm = TRUE)) #, na.rm = TRUE)
    
    # Create common time breaks
    time_breaks <- seq(min_start_time, min_end_time, interval)
  #  time_breaks <- seq(1, floor(min_end_time))
    
   # time_breaks <- seq(0, 300, interval)
    
    # Align heart rate data
    hr_aligned <- hr_sub %>%
      ungroup() %>% 
      mutate(interval = cut(heart.time, breaks = time_breaks)) %>%
      group_by(id, period, condition, interval) %>%
      summarise(HR = mean(heart.rate, na.rm = TRUE))
    
    # Align respiration data
    rsp_aligned <- rsp_sub %>%
      ungroup() %>% 
      mutate(interval = cut(rsp.pSec, breaks = time_breaks)) %>%
      group_by(id, period, condition, interval) %>%
      summarise(RR = mean(rsp.rate, na.rm = TRUE), rsp_pVal = mean(rsp.pVal, na.rm = TRUE))
    
    # Align blood pressure data
    bp_aligned <- bp_sub %>%
      ungroup() %>% 
      mutate(interval = cut(bptimes.pSec, breaks = time_breaks)) %>%
      group_by(id, period, condition, interval) %>%
      summarise(BP = mean(bp.pVal, na.rm = TRUE))
    
    # Combine aligned data frames
    aligned_sub <- full_join(hr_aligned, rsp_aligned, by = c("id", "period", "condition", "interval")) %>%
      full_join(bp_aligned, by = c("id", "period", "condition", "interval"))
    
    # Append the aligned data frame to the list
    #aligned_data[[paste(subject, period, sep = "_")]] <- aligned_sub
    aligned_data <- rbind(aligned_data, aligned_sub)
  }
}

# Combine all aligned data frames into a single data frame
#aligned_df <- bind_rows(aligned_data)
#aligned_df <- aligned_data



comb_df <- aligned_data 

#write_csv(aligned_data, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data_processed/aligned_comb_20240610.csv")

# Filter out the periods/ids with missing data
# missing_data <- as.data.frame(missing_data)
# 
# for (i in 1:dim(missing_data)[1]) {
#   comb_df_filtered <- comb_df_filtered %>% 
#     filter(!(id == missing_data$id[i] & period == missing_data$period[i]))
# }
# 
# comb_df_filtered <- comb_df_filtered %>% 
#   filter(!(id %in% c("mi003", "mi004", "mi047")))
# 
# #comb_df_filtered <- comb_df_filtered %>% 
# #  filter(condition != "MELAS")
# 
# comb_df <- comb_df_filtered 
# 
# comb_df <- comb_df_filtered %>% 
#   mutate(period = factor(period, levels = periods_str)) #%>% 
# #  unique()
# 
# comb_df_start <- comb_df
# 
# View(comb_df %>% filter(period == "CP", id == "mi005"))

```


#### data cleaning
```{r}


comb_df <- comb_df %>% 
            mutate(condition = factor(condition, levels = c("Control", "Mutation", "Deletion"))) %>% 
            mutate(period_name = period) %>% 
            mutate(period = factor(period_name, levels=c('1','2','3','4','5','6','7','8','9','10','11','12','DBT','ST','SST', 'CP'))) %>%
            mutate(period_longer = case_match(period,
                                              "1" ~ "Base",
                                              "2" ~ "Pre",
                                              c("3", "4", "5") ~ "Task",
                                              c("6", "7", "8", 
                                                "9", "10", "11", "12") ~ "Post",
                                              .default = as.character(period))) %>%
            # arrange(id, period) %>% 
            # mutate(interval = as.numeric(interval)) %>%  # Ensure interval is numeric
            arrange(id, period) %>% 
            #         case_when(
            #           period %in% c("Instruct", "PrepTask", "Task") ~ as.numeric(factor(period, levels = c("Instruct", "PrepTask", "Task"))),
            #           TRUE ~ extract_numeric(as.character(period))
            #         ),
            #         interval) %>%
            # ungroup() %>% 
            group_by(id, period_longer) %>%
            mutate(interval_longer = row_number()) %>%
            #mutate(interval_longer = as.integer(interval)) %>%
            ungroup() %>% 
            mutate(period_longer = factor(period_longer, levels = c("Base", "Pre", "Task", 
                            "Post", "DBT", "ST", "SST", "CP")))



# Plot
ggplot(comb_df %>% 
         filter(id == "mi010", period_longer == "Base") %>%
         pivot_longer(cols = c("HR", "RR", "BP"), names_to = "measurement", values_to = "value"), aes(x = interval, y = value, group = measurement, color = measurement)) +
  geom_point() +
  facet_wrap(~ period_longer) + 
  ggtitle("HR, BP, RR")

# comb_df <- comb_df %>% 
#   mutate(period_name = period) %>% 
#   mutate(period_number = case_match(
#                 period,
#                  "Base" ~ "1",
#                  "Pre" ~ "2",
#                  "Instruct" ~ "3",
#                  "PrepTask" ~ "4",
#                  "Task" ~ "5",
#                  "Post5" ~ "6",
#                  "Post10" ~ "7",
#                  "Post20" ~ "8",
#                  "Post30" ~ "9",
#                  "Post60" ~ "10",
#                  "Post90" ~ "11",
#                  "Post120" ~ "12",
#                  # "DBT",
#                  # "ST",
#                  # "SST",
#                  # "CP",
#                  .default = as.character(period))) %>% 
#             mutate(period = factor(period_name, levels = periods_str)) %>%
#             arrange(id, period) %>% 
#             mutate(period_longer = case_match(period,
#                                               "Base" ~ "Base",
#                                               "Pre" ~ "Pre",
#                                               c("Instruct", "PrepTask", "Task") ~ "Task",
#                                               c("Post5", "Post10", "Post20", 
#                                                 "Post30", "Post60", "Post90", "Post120") ~ "Post",
#                                               .default = as.character(period))) %>%
#             mutate(interval = as.numeric(interval)) %>%  # Ensure interval is numeric
#             arrange(id, period_longer,
#                     case_when(
#                       period %in% c("Instruct", "PrepTask", "Task") ~ as.numeric(factor(period, levels = c("Instruct", "PrepTask", "Task"))),
#                       TRUE ~ extract_numeric(as.character(period))
#                     ),
#                     interval) %>%
#             ungroup() %>% 
#             group_by(id, period_longer) %>%
#             mutate(interval_longer = row_number()) %>%
#             #mutate(interval_longer = as.integer(interval)) %>%
#             ungroup() %>% 
#             mutate(period_longer = factor(period_longer, levels = c("Base", "Pre", "Task", 
#                             "Post", "DBT", "ST", "SST", "CP")))

comb_df$period <- factor(comb_df$period, levels=c('1','2','3','4','5','6','7','8','9','10','11','12','DBT','ST', 'SST', "CP"))

# comb_df = left_join((comb_df %>% filter(id %in% id_rsp)), rsp_cleaned, by = c("id", "period", "interval", "condition"))  %>% 
# #  mutate(period = as.factor(period)) %>%
#   arrange(id, period, interval) %>%
#   select(-diff) %>% 
#   filter(period != "C3") 
# #%>% filter(period %in% c("CP", "SST", "ST", "DBT"))

unique(comb_df$period)

comb_df$condition <- factor(comb_df$condition, levels=c('Control', 'Mutation', 'Deletion'))
unique(comb_df$condition)
```

```{r}

#ADDED

comb_df <- comb_df %>% 
  mutate(rate_mean = HR) %>% 
  mutate(pval_mean = BP) %>% 
  mutate(rate = RR)

```


```{r}
#comb_df = comb_df %>% filter(id != "mi026" & id != "mi035"& id != "mi039" & id!= "mi029")

library(plotly)

p_hr = ggplot(comb_df, aes(x = interval, y = rate_mean, group = id, color = condition)) +
 # geom_line() +
  geom_smooth(se = FALSE, size = 0.2) +
  facet_wrap(~ period) + ggtitle("Heart Rate")
ggplotly(p_hr, mode = "markers", color = ~state, colors = "Set1")

p_bp = ggplot(comb_df, aes(x = interval, y = pval_mean, group = id, color = condition)) + 
 #  geom_line() +
  geom_smooth(se = FALSE, size = 0.2) +
  facet_wrap(~ period) + ggtitle("Blood Pressure")
plotly::ggplotly(p_bp, mode = "markers", color = ~state, colors = "Set1")
# 8 and 12 are bad, and 19 and 23

p_rsp = ggplot(comb_df, aes(x = interval, y = rate, group = id, color = condition)) + 
  geom_smooth(se = FALSE, size = 0.2) +
  facet_wrap(~ period) + ggtitle("Respiration Rate")
ggplotly(p_rsp, mode = "markers", color = ~state, colors = "Set1")

p_hr; p_bp; p_rsp

```



```{r viz_so_far_2}

# p_hr = ggplot(comb_df, aes(x = interval, y = HR, group = id, color = condition)) + 
#  # geom_smooth(se = FALSE, size = 0.2) +
#   geom_line() +
#   facet_wrap(~ period) + ggtitle("Heart Rate")
# ggplotly(p_hr, mode = "markers", color = ~condition, colors = "Set1")

ggplot(comb_df, aes(x = interval, y = HR, group = id, color = condition)) +
  geom_line() +
  facet_wrap(~ period) + 
  ggtitle("Heart Rate")

ggplot(comb_df, aes(x = interval, y = RR, group = id, color = condition)) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Respiration Rate")

ggplot(comb_df, aes(x = interval, y = BP, group = id, color = condition)) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Blood Pressure")

```

Now we have a df with all three values matched to the same intervals. However, there are some NAs because (particularly for the respiration data) sometimes the space between the data points is longer than the mapped interval. Additionally, there are some values that are unreasonable and not biologically plausible. We will filter and then interpolate.


```{r}
## remove outliers patients then filling missing values
## remove patients with no observation of particular biomarker

# rename some terms
# comb_df <- comb_df %>% 
#   rename(rate_mean = HR, pval_mean = BP, rate = RR)

# ## remove outliers observation for individuals
# for (i in 1:nrow(comb_df)) {
#   if (comb_df$rate_mean[i] < 10| comb_df$rate_mean[i] > 350| is.na(comb_df$rate_mean[i]))
#     {comb_df$rate_mean[i] = NA}
#   if (comb_df$rate[i] < 1| comb_df$rate[i] > 100| is.na(comb_df$rate[i])) 
#     {comb_df$rate[i] = NA}
#   if (comb_df$pval_mean[i] < 0| comb_df$pval_mean[i] > 500| is.na(comb_df$pval_mean[i]))
#     {comb_df$pval_mean[i] = NA}
#   
#   print(paste(comb_df$id[i], comb_df$period[i]))
#   
# }

unique(comb_df$period)
unique(comb_df$id)

comb_df <- comb_df %>% 
  mutate(rate_mean = ifelse(rate_mean > 350 | rate_mean < 10, NA, rate_mean)) %>% #Replace with NA if HR is > 150 or < 25 
  mutate(rate = ifelse(rate > 100 | rate < 1, NA, rate)) %>% 
  mutate(pval_mean = ifelse(pval_mean > 500 | pval_mean < 0, NA, pval_mean))




boundary = comb_df %>% group_by(id, period) %>%
  mutate(upper_hr = max(rate_mean, na.rm = TRUE), lower_hr = min(rate_mean, na.rm = TRUE),
         upper_bp = max(pval_mean, na.rm = TRUE), lower_bp = min(pval_mean, na.rm = TRUE),
         upper_rsp = max(rate, na.rm = TRUE), lower_rsp = min(rate, na.rm = TRUE)) %>%
  select(id, period, upper_hr, lower_hr, upper_bp, lower_bp, upper_rsp, lower_rsp) %>%
  distinct()

boundary_2 = comb_df %>% group_by(id) %>%  
  mutate(upper_hr = max(rate_mean, na.rm = TRUE), lower_hr = min(rate_mean, na.rm = TRUE),
         upper_bp = max(pval_mean, na.rm = TRUE), lower_bp = min(pval_mean, na.rm = TRUE),
         upper_rsp = max(rate, na.rm = TRUE), lower_rsp = min(rate, na.rm = TRUE)) %>%
  select(id, upper_hr, lower_hr, upper_bp, lower_bp, upper_rsp, lower_rsp) %>%
  distinct()

boundary <- boundary_2
```


```{r}
# h = 1
# 
# period_vector = unique(comb_df$period)
# library(imputeTS)
# 
# for (i in 1:length(unique(comb_df$id))) {
#   
#   cur_id = as.character(unique(comb_df$id)[i])
#   cur = comb_df %>% filter(id == cur_id)
#   
#   #for (j in c("CP", "DBT", "SST", "ST")) {
#   for (j in period_vector) {
#     
#     n = nrow(cur %>% filter(period == j))
#     
#     ## just for fixing the form, some missing value are not shown as NA, we need them to been shown as NA
#     if (nrow(cur %>% filter(period == j)) == 0) {} else if 
#     (sum(is.na((cur %>% filter(period == j))$rate_mean)) == n |
#      sum(is.na((cur %>% filter(period == j))$pval_mean)) == n |
#      sum(is.na((cur %>% filter(period == j))$pVal)) == n |
#      sum(is.na((cur %>% filter(period == j))$rate)) == n) {
#       
#       comb_df$rate_mean[h:(n+h-1)] = rep(NA, n)
#       comb_df$pval_mean[h:(n+h-1)] = rep(NA, n)
#       comb_df$pVal[h:(n+h-1)] = rep(NA, n)
#       comb_df$rate[h:(n+h-1)] = rep(NA, n)
#       h = h + n
#       
#     } else {
#       
#       cur_rate_mean = na_interpolation((cur %>% filter(period == j))$rate_mean, option = "spline")
#       cur_pval_mean = na_interpolation((cur %>% filter(period == j))$pval_mean, option = "spline")
#       cur_pval_rsp = na_interpolation((cur %>% filter(period == j))$pVal, option = "spline")
#       cur_rate_rsp = na_interpolation((cur %>% filter(period == j))$rate, option = "spline")
#       
#       comb_df$rate_mean[h:(n+h-1)] = cur_rate_mean
#       comb_df$pval_mean[h:(n+h-1)] = cur_pval_mean
#       comb_df$pVal[h:(n+h-1)] = cur_pval_rsp
#       comb_df$rate[h:(n+h-1)] = cur_rate_rsp
#     
#       h = h + n
#     }
#   }
# }
# 
# #comb_df_pre_iterpolation_2 <- comb_df
# comb_df <- comb_df_pre_iterpolation_2 


comb_df <- comb_df %>% 
  group_by(id, period) %>% 
  mutate(rate_mean = ifelse(sum(!is.na(rate_mean)) >= 2, imputeTS::na_interpolation(rate_mean, option = "spline"), rate_mean)) %>% 
  mutate(pval_mean = ifelse(sum(!is.na(pval_mean)) >= 2, imputeTS::na_interpolation(pval_mean, option = "spline"), pval_mean)) %>% 
  mutate(rate = ifelse(sum(!is.na(rate)) >= 2, imputeTS::na_interpolation(rate, option = "spline"), rate)) %>% 
  ungroup()

# 
# comb_df = comb_df %>% drop_na() %>%
#   filter(id != "mi030" & id != "mi048" & id != "mi050" & id != "mi054" & id != "mi058"& id != "mi059")
```


```{r}



```



```{r}
# h = 1
# for (i in 1:length(unique(comb_df$id))) {
#   
#   cur_id = as.character(unique(comb_df$id)[i])
#   
#   #for (j in c("CP", "DBT", "SST", "ST")) {
#   for (j in period_vector) {
#     cur = comb_df %>% filter(id == cur_id, period == j)
#     n = nrow(cur)
#     
#     if (n > 0) {  # Check if cur is not empty
#     
#     hr_upper = (boundary %>% filter(id == cur_id, period == j))$upper_hr
#     hr_lower = (boundary %>% filter(id == cur_id, period == j))$lower_hr
#     hr = ifelse(cur$rate_mean < hr_upper, cur$rate_mean, hr_upper)
#     hr = ifelse(hr > hr_lower, hr, hr_lower)
#     
#     bp_upper = (boundary %>% filter(id == cur_id, period == j))$upper_bp
#     bp_lower = (boundary %>% filter(id == cur_id, period == j))$lower_bp
#     bp = ifelse(cur$pval_mean < bp_upper, cur$pval_mean, bp_upper)
#     bp = ifelse(bp > bp_lower, bp, bp_lower)
#     
#     rsp_upper = (boundary %>% filter(id == cur_id, period == j))$upper_rsp
#     rsp_lower = (boundary %>% filter(id == cur_id, period == j))$lower_rsp
#     rsp = ifelse(cur$rate < rsp_upper, cur$rate, rsp_upper)
#     rsp = ifelse(rsp > rsp_lower, rsp, rsp_lower)
#     
#     comb_df$rate_mean[h:(n+h-1)] = hr
#     comb_df$pval_mean[h:(n+h-1)] = bp
#     comb_df$rate[h:(n+h-1)] = rsp
#     
#     h = h + n
#     }
#   }
# }
# 
# comb_df_interpolated <- comb_df
# 
# 
# h = 1
# for (i in unique(comb_df$id)) {
#   
#   cur_id <- as.character(i)
#   
#   for (j in period_vector) {
#     cur <- comb_df %>% filter(id == cur_id, period == j)
#     n <- nrow(cur)
#     
#     if (n > 0) {
#       boundary_subset <- boundary %>% filter(id == cur_id, period == j)
#       hr_upper <- boundary_subset$upper_hr
#       hr_lower <- boundary_subset$lower_hr
#       hr <- ifelse(cur$rate_mean < hr_upper, cur$rate_mean, hr_upper)
#       hr <- ifelse(hr > hr_lower, hr, hr_lower)
#       
#       bp_upper <- boundary_subset$upper_bp
#       bp_lower <- boundary_subset$lower_bp
#       bp <- ifelse(cur$pval_mean < bp_upper, cur$pval_mean, bp_upper)
#       bp <- ifelse(bp > bp_lower, bp, bp_lower)
#       
#       rsp_upper <- boundary_subset$upper_rsp
#       rsp_lower <- boundary_subset$lower_rsp
#       rsp <- ifelse(cur$rate < rsp_upper, cur$rate, rsp_upper)
#       rsp <- ifelse(rsp > rsp_lower, rsp, rsp_lower)
#       
#       index <- which(comb_df$id == cur_id & comb_df$period == j)
#       comb_df$rate_mean[index] <- hr
#       comb_df$pval_mean[index] <- bp
#       comb_df$rate[index] <- rsp
#     }
#   }
# }
# 
# comb_df_beforefilter <- comb_df_interpolated
# 
# comb_df_interpolated <- comb_df_beforefilter
# comb_df_interpolated <- comb_df_interpolated %>% filter(id != "mi001") #id != "mi059" & 
# comb_df_interpolated <- comb_df_interpolated %>% filter(id != "mi002")
# 
# ggplotly(ggplot(comb_df_interpolated, aes(x = interval, y = rate_mean, group = id, color = id)) + 
# #  geom_point(size = 0.2, alpha = 0.2) +
#   geom_line() +
#   facet_wrap(~ period) )
# 
# ggplotly(ggplot(comb_df_interpolated, aes(x = interval, y = pval_mean, group = id, color = id)) + 
# #  geom_point(size = 0.2, alpha = 0.2) +
#   geom_line() +
#   facet_wrap(~ period) )
# 
# ggplotly(ggplot(comb_df_interpolated, aes(x = interval, y = rate, group = id, color = id)) + 
# #  geom_point(size = 0.2, alpha = 0.2) +
#   geom_line() +
#   facet_wrap(~ period) )
# 
# 
# comb_df_poolside <- comb_df_interpolated
# 
# comb_df_renamed <- comb_df_poolside %>% 
#   mutate(HR = rate_mean) %>% 
#   mutate(BP = pVal) %>% 
#   mutate(RR = rate) %>% 
#   mutate(condition = case_match(condition,
#                                 "Control" ~ "Control",
#                                 "Mutation" ~ "Mutation",
#                                 "Deletion" ~ "Deletion"))
# 
# table(comb_df_renamed$condition)

# write_csv(comb_df_renamed, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/comb_df_renamed_0329.csv")




comb_df <- comb_df %>% 
  left_join(boundary, by = c("id")) %>%
  mutate(rate_mean = case_when(
    rate_mean < upper_hr & rate_mean > lower_hr ~ rate_mean,
    rate_mean >= upper_hr ~ upper_hr,
    rate_mean <= lower_hr ~ lower_hr,
    TRUE ~ rate_mean
  )) %>% 
  mutate(rate = case_when(
    rate < upper_rsp & rate > lower_rsp ~ rate,
    rate >= upper_rsp ~ upper_rsp,
    rate <= lower_rsp ~ lower_rsp,
    TRUE ~ rate
  )) %>% mutate(pval_mean = case_when(
    pval_mean < upper_bp & pval_mean > lower_bp ~ pval_mean,
    pval_mean >= upper_bp ~ upper_bp,
    pval_mean <= lower_bp ~ lower_bp,
    TRUE ~ pval_mean
  )) 
# 
# 
# 
# 
# comb_df1 <- comb_df %>% 
#   left_join(boundary, by = c("id", "period")) %>%
#   mutate(HR = case_when(
#     HR < upper_hr & HR > lower_hr ~ HR,
#     HR >= upper_hr ~ upper_hr,
#     HR <= lower_hr ~ lower_hr,
#     TRUE ~ HR
#   )) %>% 
#   mutate(RR = case_when(
#     RR < upper_rsp & RR > lower_rsp ~ RR,
#     RR >= upper_rsp ~ upper_rsp,
#     RR <= lower_rsp ~ lower_rsp,
#     TRUE ~ RR
#   )) %>% mutate(BP = case_when(
#     BP < upper_bp & BP > lower_bp ~ HR,
#     BP >= upper_bp ~ upper_bp,
#     BP <= lower_bp ~ lower_bp,
#     TRUE ~ BP
#   ))



```

```{r}
p_hr = ggplot(comb_df1 %>% filter (id == "mi010"), aes(x = interval_longer, y = rate_mean, group = id, color = condition)) + 
  #geom_line() +
  geom_point(size = 0.2, alpha = 0.2) +
  facet_wrap(~ period) 

p_bp = ggplot(comb_df, aes(x = interval, y = pval_mean, group = id, color = condition)) + 
  geom_point(size = 0.2, alpha = 0.2) +
  facet_wrap(~ period) 

p_rsp = ggplot(comb_df, aes(x = interval, y = rate, group = id, color = condition)) + 
  geom_point(size = 0.2, alpha = 0.2) +
  facet_wrap(~ period) 
a = 1
p_hr; p_bp; p_rsp
```

# TE calculation

```{r}

comb_df_pre_te <- comb_df

# add differences variables

comb_df <- comb_df %>%
  mutate(rate_diff = HR - lag(HR),
         pval_diff = BP - lag(BP),
         pval_rsp_diff = rsp_pVal - lag(rsp_pVal),
         rate_rsp_diff = RR - lag(RR)) %>%
  rename(rate_hr = HR, 
    rate_hr_diff = rate_diff,
         pval_bp = BP, 
    pval_bp_diff = pval_diff,
         pval_rsp = rsp_pVal, rate_rsp = RR) 



# set comb_df, periods_str

library(RTransferEntropy)
library(future)

#cur <- comb_df %>% filter(id == "mi069", period == "Post")

future::plan(multisession)

# Initialize data frames to store results
te_df <- data.frame()
skipped_te <- data.frame()


# Loop through ids and periods (this takes a while!)

#for(id_str in ids %>% pull(id)) {
for(id_str in subject_ids) {
# for(period_str in periods_str){
for(period_str in unique(comb_df$period_longer)){

#for(period_str in c("Post", "SST", "ST")){

# 
# for(id_str in c("mi021", "mi053", "mi069", "mi100", "mi053", "mi037")) {
#        for(period_str in c("Post60", "PrepTask", "Post5")){
         
  #  cur = comb_df %>% filter(id == id_str, period == period_str)
    cur = comb_df %>% filter(id == id_str, period_longer == period_str) %>% mutate(period = period_longer)
    
    print(id_str)
    print(period_str)
    
    # Skip if empty
    if (dim(cur)[1] == 0) {
      skipped_instance <- data.frame(skipped_id = id_str,
                                     skipped_period = period_str,
                                     reason = "null")
      skipped_te <- rbind(skipped_te, skipped_instance)
      next
    }
    
    # Calculate TE within tryCatch
    tryCatch({
      
      # TE on differences
      TE1hb_diff = transfer_entropy(cur$rate_hr_diff, cur$pval_bp_diff, type = 'quantiles', lx = 1, ly = 1)#, nboot = 0)#, burn = 20)
      TE2hr_diff = transfer_entropy(cur$rate_hr_diff, cur$rate_rsp_diff, type = 'quantiles', lx = 1, ly = 1)#, nboot = 0)#, burn = 20)
      TE3br_diff = transfer_entropy(cur$pval_bp_diff, cur$rate_rsp_diff, type = 'quantiles', lx = 1, ly = 1)#, nboot = 0)#, burn = 20)
      
      # TE on original time series
      TE1hb = transfer_entropy(cur$rate_hr, cur$pval_bp, type = 'quantiles', lx = 1, ly = 1)#, burn = 20)
      TE2hr = transfer_entropy(cur$rate_hr, cur$rate_rsp, type = 'quantiles', lx = 1, ly = 1)#, burn = 20)
      TE3br = transfer_entropy(cur$pval_bp, cur$rate_rsp, type = 'quantiles', lx = 1, ly = 1)#, burn = 20)
      
      te_row <- data.frame(
        id = id_str,
        period = period_str,
      # condition = condition,  
        te_hr_bp_diff = TE1hb_diff$coef[1, 1],
        te_bp_hr_diff = TE1hb_diff$coef[2, 1],
        te_hr_rr_diff = TE2hr_diff$coef[1, 1],
        te_rr_hr_diff = TE2hr_diff$coef[2, 1],
        te_bp_rr_diff = TE3br_diff$coef[1, 1],
        te_rr_bp_diff = TE3br_diff$coef[2, 1],
        te_hr_bp = TE1hb$coef[1, 1],
        te_bp_hr = TE1hb$coef[2, 1],
        te_hr_rr = TE2hr$coef[1, 1],
        te_rr_hr = TE2hr$coef[2, 1],
        te_bp_rr = TE3br$coef[1, 1],
        te_rr_bp = TE3br$coef[2, 1]#,
        # te_hr_bp_diff_p = TE1hb_diff$coef[1, 4],
        # te_bp_hr_diff_p = TE1hb_diff$coef[2, 4],
        # te_hr_rr_diff_p = TE2hr_diff$coef[1, 4],
        # te_rr_hr_diff_p = TE2hr_diff$coef[2, 4],
        # te_bp_rr_diff_p = TE3br_diff$coef[1, 4],
        # te_rr_bp_diff_p = TE3br_diff$coef[2, 4]#,
        # te_hr_bp_p = TE1hb$coef[1, 4],
        # te_bp_hr_p = TE1hb$coef[2, 4],
        # te_hr_rr_p = TE2hr$coef[1, 4],
        # te_rr_hr_p = TE2hr$coef[2, 4],
        # te_bp_rr_p = TE3br$coef[1, 4],
        # te_rr_bp_p = TE3br$coef[2, 4]
      )
      
      te_df <- rbind(te_df, te_row)
      
    }, error = function(e) {
      # Handle the error by storing the skipped instance
      skipped_instance <- data.frame(skipped_id = id_str,
                                     skipped_period = period_str,
                                     reason = "error")
      skipped_te <- rbind(skipped_te, skipped_instance)
      
      # Print error message for debugging
      print(paste("Error for id:", id_str, "and period:", period_str))
      print(e)
    })
  }
}


print(te_df)
print(skipped_te)

#colnames(te_df) <- names(te_row)

# add condition back in 
id_cond <- comb_df %>% 
  ungroup() %>% 
  select(id, condition) %>% 
  unique()

te_df <- left_join(te_df, id_cond, by = "id") %>% 
  select(id, period, condition, everything())

plan(sequential)
# 
# write_csv(te_df, file = file.path("./data_processed", "te_df_notdetrended_20240524.csv"))
# write_csv(skipped_te, file = file.path("./data_processed", "skipped_te_notdetrended_20240524.csv"))

te_df_8 <- te_df
#write_csv(te_df, file = file.path("./data_processed", "te_df_8_notdetrended_20240524.csv"))
#write_csv(skipped_te, file = file.path("./data_processed", "skipped_te_8_notdetrended_20240524.csv"))

te_df_l_1 <- te_df_8
te_df_l_3 <- te_df

```

Many were skipped, let's explore that. 

```{r te_completeness}

# Checking completeness
te_df %>% 
  group_by(id) %>% 
  summarize(n_periods = n()) %>% 
  group_by(n_periods) %>% 
  summarize(n = n())

te_df_original %>% 
  group_by(id) %>% 
  summarize(n_periods = n()) %>% 
  group_by(n_periods) %>% 
  summarize(n = n())

te_df %>% 
  group_by(period) %>% 
  summarize(n = n())

te_df_original %>% 
  na.omit() %>% 
  group_by(phase) %>%
  summarize(n = n())

#   n_periods     n
#       <int> <int>
# 1         5     1
# 2         9     1
# 3        10     1
# 4        11     3
# 5        12     4
# 6        13     5
# 7        14    25
# 8        15    50
# 9        16    15

# Checking completeness
te_df %>% 
  group_by(id) %>% 
  filter(period == "PrepTask") %>% 
  summarize(n_periods = n()) %>% 
#  filter(n_periods == 15) %>% 
  summarize(n = n())

# 34 ----> many,  many PrepTasks failed

te_df %>% 
  group_by(id, period) %>% 
  summarize(n = n())
  
  
#te_df_0 <- read_csv(file.path("./data_processed", "te_df_notdetrended_20240524.csv"))


```

# Comparing TE values (first pass)

# T tests and Wilcoxon Signed Rank tests (Mann-Whitney)

```{r}

#head(all)
library(tidyr)
library(broom)

# Assuming df has columns: id, condition, period, and metric

combined_summary <- te_df

#all_metric_names <- colnames(combined_summary %>% #select("te_hr_bp_diff":"te_rr_bp_diff"))
all_metric_names <- colnames(combined_summary %>% select("te_hr_bp_diff":"te_rr_bp"))


 
# combined_summary <- te_df_original %>% 
#   select(id, Direction, Status, TE_value, phase) %>% 
#   group_by(id, Status, phase) %>% 
#   pivot_wider(names_from = Direction,
#               values_from = TE_value) %>% 
#   rename(condition = Status) %>% 
#   rename(period = phase)
# all_metric_names <- colnames(combined_summary %>% ungroup() %>% select("Hr_to_Bp":"Rsp_to_Bp"))



# all_metric_names <- colnames(combined_summary %>% 
#                                select(-"id", -"condition", -"period"))

# Assuming df has columns: id, condition, period, and multiple metrics
# all_metric_names is your vector of metric names
library(dplyr)
library(tidyr)

results_list <- lapply(all_metric_names, function(metric) {
  metric_data <- combined_summary %>%
    select(id, condition, period, !!sym(metric)) %>%
    rename(metric_value = !!sym(metric))
  
  pairwise_comparisons <- metric_data %>%
    group_by(period) %>%
    do({
      subset_data <- .
      comparisons <- expand.grid(unique(subset_data$condition), unique(subset_data$condition))
      comparisons <- comparisons[comparisons$Var1 != comparisons$Var2,]
      
      # Initialize results dataframe
      results <- data.frame()
      
      for (i in 1:nrow(comparisons)) {
        cond1 <- comparisons$Var1[i]
        cond2 <- comparisons$Var2[i]
        
        data1 <- subset(subset_data, condition == cond1)$metric_value
        data2 <- subset(subset_data, condition == cond2)$metric_value
        
        if (length(data1) > 1 && length(data2) > 1) {
          t_test_result <- t.test(data1, data2)
          wilcoxon_test_result <- wilcox.test(data1, data2, exact = FALSE)  # allow approximate p-value
          
          results <- rbind(results, data.frame(
            Var1 = cond1,
            Var2 = cond2,
            period = unique(subset_data$period),
            mean_1 = t_test_result$estimate[1],
            mean_2 = t_test_result$estimate[2],
            n_1 = length(data1),
            n_2 = length(data2),
            t_test_p = t_test_result$p.value,
            wilcoxon_test_p = wilcoxon_test_result$p.value,
            sign_g1_v_g2 = if_else(sign(t_test_result$estimate[1] - t_test_result$estimate[2]) == 1, paste0(cond1, " higher than ", cond2), paste0(cond1, " lower than ", cond2))
          ))
        }
      }
      results
    }) %>%
    unnest(cols = everything())  # Specify columns to unnest
  
  group_means <- metric_data %>%
    group_by(condition, period) %>%
    summarise(mean_metric = mean(metric_value, na.rm = TRUE), .groups = 'drop')
  
  # Combine results
  results <- left_join(pairwise_comparisons, group_means, by = c("Var1" = "condition", "period"))
  results$metric <- metric  # Add the metric name to the results
  
  return(results)
})

results_list <- bind_rows(results_list)  # Combine all results into one data frame

results_list <- results_list %>% 
  filter(Var1 != "MELAS" & Var2 != "MELAS") %>% 
  filter(!(Var1 == "Control" & Var2 == "Mutation")) %>% 
  filter(!(Var1 == "Control" & Var2 == "Deletion")) %>% 
  filter(!(Var1 == "Deletion" & Var2 == "Mutation")) %>% 
  rename(Group1 = Var1) %>% 
  rename(Group2 = Var2) %>% 
  select(Group1, Group2, period, metric, sign_g1_v_g2, t_test_p, wilcoxon_test_p, everything())

p_val_per_period <- results_list

results_t_sig <- results_list %>% 
  filter(t_test_p <= 0.15)

results_wilcoxon_sig <- results_list %>% 
  filter(wilcoxon_test_p <= 0.15)

# write_csv(p_val_per_period, file = file.path("./data_processed", "p_val_per_period_8_notdetrended_108_20240603.csv"))

```


