---
title: "Map Intervals and Combine Metrics"
author: "Christine Lucille Kuryla"
date: "2024-05-22"
output: html_document
---


For sensitivity analyses, try to vary the length of the mapping time interval (currently 4 seconds), the interpolation method (currently spline, natural), the limits of the outlier removal, and setting the removed outliers to be the boundary values instead of interpolated.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Load data

```{r load_data}

hr_raw <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/raw_data_110_202404/hr_raw_108.csv")
rr_raw <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/raw_data_110_202404/rr_raw_105.csv")
bp_raw <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/raw_data_110_202404/bp_raw_107.csv")

# for graphing
# string vector
periods_str <- c("Base",
                 "Pre",
                 "Instruct",
                 "PrepTask",
                 "Task",
                 "Post5",
                 "Post10",
                 "Post20",
                 "Post30",
                 "Post60",
                 "Post90",
                 "Post120",
                 "DBT",
                 "ST",
                 "SST",
                 "CP")

# factor vector
# create a vector with the periods as factors in the correct order (not alphabetical)
# this is for ease of plotting later
periods_factor <- factor(periods_str, levels = periods_str)

```

Hard code missing data

```{r missing_data}

# See import RMD for details
# For now, if there exists at least one missing modality, we will remove that id/period combo.

missing_data <- (rbind(
  c("mi002", "PrepTask"), #rsp 
  c("mi002", "Task"), #rsp
c("mi009", "Base"), #rttimes
c("mi012", "Instruct"), #bptimes
c("mi012", "Post5"), #bptimes
c("mi012", "Post120"), #bptimes
c("mi012", "DBT"), #bptimes
c("mi012", "ST"), #bptimes
c("mi016", "Instruct"), # rttimes for Instruct missing
c("mi024", "Post120"), # bptimes for Post120 missing
c("mi028", "SST"), # rttimes for SST missing
c("mi032", "Post10"), # rsp for Post10 missing
c("mi041", "Post120"), # mi041Post120_bptimes.csv 
c("mi046", "Post5"), # mi046Post5_rttimes.csv 
c("mi050", "ST"), # mi050ST_bptimes.csv 
c("mi053", "Post60"), # mi053Post60_rsp.csv 
c("mi083", "SST"), # mi083SST_rttimes.csv , mi083SST_rsp.csv 
c("mi087", "DBT"), # mi087DBT_bptimes.csv 
c("mi088", "ST"), # mi088ST_bptimes.csv 
c("mi098", "Post30"), # mi098Post30_bptimes.csv 
c("mi105", "Task"), # mi105Task_rsp.csv 
c("mi108", "PrepTask"), # mi108PrepTask_rsp.csv 
c("mi110", "Post5"), # mi110Post5_rsp.csv , mi110Post10_rsp.csv , mi110Post120_bptimes.csv 
c("mi110", "Post10"),
c("mi110", "Post120"),
c("mi109", "CP"), # mi109 is missing rsp for CP only
c("mi070", "CP"), # mi070 is missing rttimes for CP only
c("mi026", "CP"), # mi026 is missing bptimes for CP only but there is another file mi026CP.2.bi --> check what this means
c("mi094", "CP"), # mi094 is missing the last 4 tasks (CP, SST, ST, DBT)
c("mi094", "SST"),
c("mi094", "ST"),
c("mi094", "DBT"),
c("mi104", "CP"), # mi104 is missing the last 4 tasks (CP, SST, ST, DBT)
c("mi104", "SST"),
c("mi104", "ST"),
c("mi104", "DBT"),
c("mi106", "CP"), # mi106 is missing all CP, rttimes in Task, and bptimes in DBT
c("mi106", "Task"),
c("mi106", "DBT"),
c("mi069", "Pre") # missing BP
))

colnames(missing_data) <- c("id", "period")

# also just take out mi003, mi004, mi047

# Summary

# mi003 is missing rsp
# mi004 is missing rsp
# mi047 is missing rsp and bptimes

# mi014 does not have correct files (it just says acquisition?) --> omitted
# mi075 is missing completely


```

Initial EDA 

```{r initial_eda}
# Visualize the raw data

ggplot(hr_raw, aes(x = heart.time, y = heart.rate, group = id, color = condition)) + 
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Heart Rate")

ggplot(rr_raw, aes(x = rsp.time, y = rsp.rate, group = id, color = condition)) + 
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Respiration Rate")

ggplot(bp_raw, aes(x = bp.time, y = bp.pVal, group = id, color = condition)) +
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Blood Pressure")
```

Initial outlier value removal

```{r initial_cleaning}
# Clearly there are some unreasonable values where the sensors or calculations were off. Let's replace them with NAs

## remove outliers observation for individuals

hr_df <- hr_raw
rr_df <- rr_raw
bp_df <- bp_raw

# Reasonable ranges for heart rate
for (i in 1:nrow(hr_df)){
  if (hr_df$heart.rate[i] < 10| hr_df$heart.rate[i] > 300| is.na(hr_df$heart.rate[i]))
    {hr_df$heart.rate[i] = NA}
}

# Reasonable ranges for respiration rate
for (i in 1:nrow(rr_df)){
  if (rr_df$rsp.rate[i] < 1| rr_df$rsp.rate[i] > 150| is.na(rr_df$rsp.rate[i]))
    {rr_df$rsp.rate[i] = NA}
}

# Reasonable ranges for blood pressure
for (i in 1:nrow(bp_df)){
  if (bp_df$bp.pVal[i] < 30| bp_df$bp.pVal[i] > 350| is.na(bp_df$bp.pVal[i]))
    {bp_df$bp.pVal[i] = NA}
}
```


```{r initial_cleaning_visualization}
ggplot(hr_df, aes(x = heart.time, y = heart.rate, group = id, color = condition)) + 
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Heart Rate")

ggplot(rr_df, aes(x = rsp.time, y = rsp.rate, group = id, color = condition)) + 
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Respiration Rate")

ggplot(bp_df, aes(x = bp.time, y = bp.pVal, group = id, color = condition)) + 
 # geom_smooth(se = FALSE, size = 0.2) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Blood Pressure")

```


Map data to matching intervals

```{r map_to_matching_intervals}
# Assuming you have the following data frames:
# - hr_df: heart rate data with columns 'id', 'period', 'time', 'rate'
# - rr_df: respiration data with columns 'id', 'period', 'pSec', 'pVal', 'rate'
# - bp_df: blood pressure data with columns 'id', 'period', 'pSec', 'pVal'

# hr_df <- hr_raw %>%
#   mutate(time = heart.time) %>%
#   mutate(rate = heart.rate) %>%
#   select(id, period, condition, time, rate)
# 
# rr_df <- rr_raw %>% 
#   mutate(pSec = rsp.pSec) %>% 
#   mutate(pVal = rsp.pVal) %>% 
#   mutate(rate = rsp.rate) %>% 
#   select(id, period, condition, pSec, pVal, rate)
# 
# bp_df <- bp_raw %>% 
#   mutate(pSec = bptimes.pSec) %>% 
#   mutate(pVal = bp.pVal) %>% 
#   select(id, period, condition, pSec, pVal)

# hr_df <- hr_raw
# rr_df <- rr_raw
# bp_df <- bp_raw
  

# Desired time interval (in seconds)
interval <- 3

# Get unique subject IDs and periods
subject_ids <- unique(c(hr_df$id, rr_df$id, bp_df$id))
periods <- unique(c(hr_df$period, rr_df$period, bp_df$period))

# Initialize an empty list to store the aligned data frames
aligned_data <- list()

# Iterate over each subject and period
for (subject in subject_ids) {
  for (period in periods) {
    # Subset data for the current subject and period
    hr_sub <- hr_df %>% filter(id == subject, period == period)
    rsp_sub <- rr_df %>% filter(id == subject, period == period)
    bp_sub <- bp_df %>% filter(id == subject, period == period)
    
    # Find the common time range
    min_start_time <- min(hr_sub$heart.time, rsp_sub$rsp.pSec, bp_sub$bptimes.pSec, na.rm = TRUE)
    max_end_time <- max(hr_sub$heart.time, rsp_sub$rsp.pSec, bp_sub$bptimes.pSec, na.rm = TRUE)
    
    # Create common time breaks
    time_breaks <- seq(min_start_time, max_end_time, interval)
    
    # Align heart rate data
    hr_aligned <- hr_sub %>%
      mutate(interval = cut(heart.time, breaks = time_breaks)) %>%
      group_by(id, period, condition, interval) %>%
      summarise(HR = mean(heart.rate, na.rm = TRUE))
    
    # Align respiration data
    rsp_aligned <- rsp_sub %>%
      mutate(interval = cut(rsp.pSec, breaks = time_breaks)) %>%
      group_by(id, period, condition, interval) %>%
      summarise(RR = mean(rsp.rate, na.rm = TRUE), rsp_pVal = mean(rsp.pVal, na.rm = TRUE))
    
    # Align blood pressure data
    bp_aligned <- bp_sub %>%
      mutate(interval = cut(bptimes.pSec, breaks = time_breaks)) %>%
      group_by(id, period, condition, interval) %>%
      summarise(BP = mean(bp.pVal, na.rm = TRUE))
    
    # Combine aligned data frames
    aligned_sub <- full_join(hr_aligned, rsp_aligned, by = c("id", "period", "condition", "interval")) %>%
      full_join(bp_aligned, by = c("id", "period", "condition", "interval"))
    
    # Append the aligned data frame to the list
    aligned_data[[paste(subject, period, sep = "_")]] <- aligned_sub
  }
}

# Combine all aligned data frames into a single data frame
aligned_df <- bind_rows(aligned_data)

# Filter out the periods/ids with missing data

comb_df_filtered <- aligned_df 

missing_data <- as.data.frame(missing_data)

for (i in 1:dim(missing_data)[1]) {
  comb_df_filtered <- comb_df_filtered %>% 
    filter(!(id == missing_data$id[i] & period == missing_data$period[i]))
}

comb_df_filtered <- comb_df_filtered %>% 
  filter(!(id %in% c("mi003", "mi004", "mi047")))

#comb_df_filtered <- comb_df_filtered %>% 
#  filter(condition != "MELAS")

comb_df <- comb_df_filtered %>% 
  mutate(period = factor(period, levels = periods_str)) %>% 
  unique()


```


```{r viz_so_far_2}

# p_hr = ggplot(comb_df, aes(x = interval, y = HR, group = id, color = condition)) + 
#  # geom_smooth(se = FALSE, size = 0.2) +
#   geom_line() +
#   facet_wrap(~ period) + ggtitle("Heart Rate")
# ggplotly(p_hr, mode = "markers", color = ~condition, colors = "Set1")

ggplot(comb_df, aes(x = interval, y = HR, group = id, color = condition)) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Heart Rate")

ggplot(comb_df, aes(x = interval, y = RR, group = id, color = condition)) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Respiration Rate")

ggplot(comb_df, aes(x = interval, y = BP, group = id, color = condition)) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Blood Pressure")

```

Now we have a df with all three values matched to the same intervals. However, there are some NAs because (particularly for the respiration data) sometimes the space between the data points is longer than the mapped interval. 

Here, we will interpolate the NAs.

```{r interpolate_nas}

library(imputeTS)

# using old code might make it prettier later

comb_df_orig <- comb_df

h = 1

period_vector = unique(comb_df$period)
library(imputeTS)

for (i in 1:length(unique(comb_df$id))) {
  
  cur_id = as.character(unique(comb_df$id)[i])
  cur = comb_df %>% filter(id == cur_id)
  
  #for (j in c("CP", "DBT", "SST", "ST")) {
  for (j in period_vector) {
    
    n = nrow(cur %>% filter(period == j))

    # just for fixing the form, some missing value are not shown as NA, we need them to been shown as NA
    if (nrow(cur %>% filter(period == j)) == 0) {} else if
    (sum(is.na((cur %>% filter(period == j))$HR)) == n |
     sum(is.na((cur %>% filter(period == j))$RR)) == n |
     sum(is.na((cur %>% filter(period == j))$rsp_pVal)) == n |
     sum(is.na((cur %>% filter(period == j))$BP)) == n) {

      comb_df$HR[h:(n+h-1)] = rep(NA, n)
      comb_df$RR[h:(n+h-1)] = rep(NA, n)
      comb_df$rsp_pVal[h:(n+h-1)] = rep(NA, n)
      comb_df$BP[h:(n+h-1)] = rep(NA, n)
      h = h + n

    } else {
      
      cur_HR = na_interpolation((cur %>% filter(period == j))$HR, option = "spline", maxgap = 3)
      cur_RR = na_interpolation((cur %>% filter(period == j))$RR, option = "spline", maxgap = 3)
      cur_rsp_pVal = na_interpolation((cur %>% filter(period == j))$rsp_pVal, option = "spline", maxgap = 3)
      cur_BP = na_interpolation((cur %>% filter(period == j))$BP, option = "spline", maxgap = 3)
      
      comb_df$HR[h:(n+h-1)] = cur_HR
      comb_df$RR[h:(n+h-1)] = cur_RR
      comb_df$rsp_pVal[h:(n+h-1)] = cur_rsp_pVal
      comb_df$BP[h:(n+h-1)] = cur_BP
    
      h = h + n
    }
  }
}

#write_csv(comb_df, file = file.path("./data_processed", "comb_df_20240524.csv"))


# remove the missingness?
#comb_df <- comb_df %>% drop_na()

```

