---
title: "Data Simulations, Tests, and Sanity Checks"
author: "Christine Lucille Kuryla"
date: "2024-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmd is a place for the simulations, tests, and sanity checks to live so they don't crowd the normal Rmds. 

# Simulate data and test TE 

1. Simulate ground truth as an equation Y which is dependent on equation X. (This has equal time intervals)
2. Reverse engineer ground truth heart rate into RR intervals
3. Calculate rate from these RR intervals as in our pipeline
4. Map to equal time intervals as in our pipeline
5. Test if TE has same result

```{r}

# t is the independent variable (in our case, time)


# Create Ground Truth

# Simulated heart rate data (assuming data from the previous simulation)
set.seed(918)
n <- 1000  # Number of time points
phi <- 0.9  # Autoregressive parameter
sigma <- 0.1  # Standard deviation of the noise

# Simulate the first time series (X) using an AR(1) process
X <- numeric(n)
X[1] <- rnorm(1, mean = 0, sd = sigma)
for (t in 2:n) {
  X[t] <- phi * X[t - 1] + rnorm(1, mean = 0, sd = sigma)
}

# Simulate the second time series (Y) as a function of past values of X
Y <- numeric(n)
for (t in 2:n) {
 # Y[t] <- 0.5 * X[t] + 0.3 * X[t - 1] + rnorm(1, mean = 0, sd = sigma)
  Y[t] <- 0.3 * X[t - 1] + rnorm(1, mean = 0, sd = sigma)
}

# Create a data frame with uniform time intervals
time_steps <- seq(from = 0, by = 1, length.out = n)
heart_rate_data <- tibble(time = time_steps, heart_rate_X = X, heart_rate_Y = Y) %>% 
  mutate(orig_x = heart_rate_X) %>% 
  mutate(heart_rate_X = 80 + 50*orig_x) %>% 
  mutate(orig_y = heart_rate_Y) %>% 
  mutate(heart_rate_Y = 80 + 50*orig_y)

# check TE
transfer_entropy(heart_rate_data$heart_rate_X, heart_rate_data$heart_rate_Y, lx = 1, ly = 1)
transfer_entropy(heart_rate_data$orig_x, heart_rate_data$orig_y, lx = 1, ly = 1)

# Save ground truth for later
ground_truth_x <- heart_rate_data %>% select(time, heart_rate_X)
ground_truth_y <- heart_rate_data %>% select(time, heart_rate_Y)


# Plot ground truth
HR_long <- heart_rate_data %>%
  pivot_longer(cols = starts_with("heart_rate_"), names_to = "series", values_to = "value")

ggplot(HR_long, aes(x = time, y = value, color = series)) +
  geom_line() +
  labs(title = "Simulated HR Time Series", x = "Time", y = "HR") +
  theme_minimal()


# Now let's see if our process works

# Convert heart rate to RR intervals
heart_rate_data <- heart_rate_data %>%
  mutate(RR_interval_X = 60 / heart_rate_X, 
         RR_interval_Y = 60 / heart_rate_Y)

# Reverse engineer the RR interval time series
RR_intervals_X <- cumsum(heart_rate_data$RR_interval_X)
RR_intervals_Y <- cumsum(heart_rate_data$RR_interval_Y)

# Create a data frame for RR intervals
RR_intervals_data <- tibble(time = time_steps, 
                            pSec_x = RR_intervals_X, 
                            pSec_y = RR_intervals_Y)

# calculate rate as in our pipeline
RR_intervals_data <- RR_intervals_data %>% 
  mutate(diff_x = pSec_x - lag(pSec_x)) %>%
  mutate(sum_t2_x = pSec_x + lag(pSec_x)) %>% 
  mutate(rate_x = 60/diff_x) %>%
  mutate(time_x = sum_t2_x/2) %>% 
  mutate(diff_y = pSec_y - lag(pSec_y)) %>%
  mutate(sum_t2_y = pSec_y + lag(pSec_y)) %>% 
  mutate(rate_y = 60/diff_y) %>%
  mutate(time_y = sum_t2_y/2)


ggplot(RR_intervals_data, aes(x = time_x, y = rate_x)) +
  geom_line() +
  labs(title = "Simulated RR Intervals Time Series X", x = "Time", y = "HR") +
  theme_minimal()

ggplot(RR_intervals_data, aes(x = time_y, y = rate_y)) +
  geom_line() +
  labs(title = "Simulated RR Intervals Time Series Y", x = "Time", y = "HR") +
  theme_minimal()

rr_calc_rate_x <- RR_intervals_data %>% select(time_x, rate_x)
rr_calc_rate_y <- RR_intervals_data %>% select(time_y, rate_y)



##### Map to constant interval

# Find the common time range
  min_start_time <- min(rr_calc_rate_x$time_x, rr_calc_rate_y$time_y, na.rm = TRUE)
  max_end_time <- max(rr_calc_rate_x$time_x, rr_calc_rate_y$time_y, na.rm = TRUE)

# desired time interval in seconds 
  
  interval <- 1
  
# Create common time breaks
 time_breaks <- seq(min_start_time, max_end_time, interval)

hr_aligned_x <- rr_calc_rate_x %>%
  mutate(interval = cut(time_x, breaks = time_breaks)) %>%
  group_by(interval) %>% 
  summarise(HR_x = mean(rate_x, na.rm = TRUE)) %>% 
  mutate(interval_x = as.integer(interval)) %>% 
  mutate(diff_x = HR_x - lag(HR_x)) 

hr_aligned_y <- rr_calc_rate_y %>%
  mutate(interval = cut(time_y, breaks = time_breaks)) %>%
  group_by(interval) %>% 
  summarise(HR_y = mean(rate_y, na.rm = TRUE)) %>% 
  mutate(interval_y = as.integer(interval)) %>% 
  mutate(diff_y = HR_y - lag(HR_y)) 

hr_aligned <- inner_join(hr_aligned_x, hr_aligned_y, by = "interval")

transfer_entropy(hr_aligned$HR_x, hr_aligned$HR_y, lx = 1, ly = 1)
transfer_entropy(hr_aligned$diff_x, hr_aligned$diff_y, lx = 1, ly = 1)




```

This gave the expected results in the normal time series but not in the differenced time series.

Let's test the NA interpolations.

# Test NA interpolation methods

```{r}

# Randomly add in NAs

hr_aligned <- hr_aligned %>% 
  mutate(hr_na_x = HR_x) %>% 
  mutate(hr_na_y = HR_y) %>% 
  mutate(hr_interpolated_x = HR_x) %>% 
  mutate(hr_interpolated_y = HR_y) 

# Generate random indicies to make NA
na_indicies_x <- sample(1:nrow(hr_aligned), size = 0.2 * nrow(hr_aligned))
na_indicies_y <- sample(1:nrow(hr_aligned), size = 0.2 * nrow(hr_aligned))

# Put NAs in heart rate data
hr_aligned$hr_na_x[na_indicies_x] <- NA
hr_aligned$hr_na_y[na_indicies_y] <- NA

# Interpolate NAs
hr_aligned <- hr_aligned %>% 
  mutate(hr_interpolated_x = na_interpolation(hr_na_x, option = "spline", maxgap = 3)) %>% 
  mutate(hr_interpolated_y = na_interpolation(hr_na_y, option = "spline", maxgap = 3))

transfer_entropy(hr_aligned$hr_interpolated_x, hr_aligned$hr_interpolated_y, lx = 1, ly = 1)
transfer_entropy(hr_aligned$diff_x, hr_aligned$diff_y, lx = 1, ly = 1)




```

