---
title: "Single Variable Measures"
author: "Christine Lucille Kuryla"
date: "2024-02-22"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    code_folding: show 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(RColorBrewer)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))


```

# Introduction

## MiSBIE Data Background

See "misbie_explore.Rmd" for more details about the variables and some EDA. 

*tl;dr:* There are three groups: healthy, mutation, and deletion. The mutation and deletion groups are subjects with genetic mitochondrial abnormalities. Heart rate, respiration rate, and blood pressure measures were recorded for 71 subjects over several tasks and recovery periods.

## Data Forms

Note the raw data allows for HR and RR to be analyzed independent of rate because the raw data is each systole/diastole beat and each inspiration/expiration, respectively. Rate can be calculated per interval or may be interpolated. 

At least two types of forms of the data will be used:

1. Non-constant delta-t --> raw data for heart rate and respiration rate
2. Constant delta-t --> interpolated, methods from transfer entropy
  a. This may be split into periods that align as well as more appropriate time periods
  b. Must be calculated and subsequently interpolated

## Objective

This Rmd will explore various complexity measures from single variables (heart rate, blood pressure, or respiration rate) over time. 
    
## Period Key
Base (period 1)
pre (period 2)
task (period 3-5)
post (period 6-12)

## Some Background and Resources

* Complexity testing techniques for time series data: A comprehensive literature review (Tang, 2015)
    * https://www.sciencedirect.com/science/article/pii/S0960077915002817

* An Overview of Heart Rate Variability Metrics and Norms
    * https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5624990/ 
    
* Nonlinear Methods Most Applied to Heart-Rate Time Series: A Review
    * https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7516766/ 
    
    
Respiratory: https://neuropsychology.github.io/NeuroKit/examples/rsp_rrv/rsp_rrv.html   

# Current thoughts

* Periods 1-6 seem to be the most uniform (same amount of time, etc). Although we may just concatenate them, I feel that may be problematic. I think I am going to calculate each measure for the 6 periods 6-12 (post) (maybe all concatenated too?) and then take the arithmetic mean and use that value. Might be a problem because this is the recovery period. This is why I'll do all of 'post' as well all six individually. 
* With EDA, period 11 seems the cleanest, so I will just use period 10 as I am working on the project, then do the average of all 6 later. 
    
# Packages

```{r load_packages, echo = FALSE, message = FALSE}
library(nonlinearTseries)
library(CGManalyzer)
#library(MSMVSampEn) # might have problems, see article

#library(Rlibeemd) 

library(plotly)
library(tidyverse)

```

# Load and process data

## Load Raw Data: Non-constant delta t (time interval)

```{r load_raw_data, eval=FALSE}

# For now, we will focus on heart rate.

hr_raw <- read_csv("./data/data_from_email/hr_cleaned.csv")
bp_raw <- read_csv("./data/data_from_email/bp_cleaned.csv")
rsp_raw <- read_csv("./data/data_from_email/rsp_cleaned.csv")
pt_cond_raw <- read_csv("./data/data_from_email/patient condition.csv")

# Sample size
length(unique(hr_raw))
length(unique(bp_raw))
length(unique(rsp_raw))

hr_raw %>%
  group_by(condition) %>%
  summarise(UniqueSubjects = n_distinct(id)) 

# Clean up the data and add column for time between beats
hr_raw_final <- hr_raw %>% 
  filter(condition %in% c("Control", "Deletion", "Mutation")) %>% 
  mutate(rate = ifelse(rate > 150 | rate < 25, NA, rate)) %>% #Replace with NA if HR is > 150 or < 25 
  mutate(delta_t  = pSec - lag(pSec)) %>% 
  mutate(delta_t = ifelse(delta_t > 5 | delta_t < -5, NA, delta_t)) # get rid of the first data point (lag) and the unreasonable data points




hr_df <- hr_raw_final





rm(hr_raw_final)

# for future use
id_cond_df <- hr_df %>% 
  select(id, condition) %>% 
  unique()

```

## Load Interpolated data: constant, measure-specific time interval

```{r}
#do this later
```


## Load Interpolated data: constant, aligned time interval (see TE)

```{r load_interpolated_data}

# saveRDS(id_to_group, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/data_20240317old/id_to_group.RDS")
# saveRDS(id_cond_df, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/data_20240317old/id_cond_df.RDS")
# saveRDS(comb_df_1to12, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/data_20240317old/comb_df_1to12.RDS")
# saveRDS(comb_df_tasks, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/data_20240317old/comb_df_tasks.RDS")

id_cond_df <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/pt_cond97.csv")
id_to_group <- id_cond_df
#id_to_group <- readRDS("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/data_20240317old/id_to_group.RDS")
comb_df_1to12 <- readRDS("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/data_20240317old/comb_df_1to12.RDS")
comb_df_tasks <- readRDS("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/data_20240317old/comb_df_tasks.RDS")

comb_df_6_12 <- comb_df_1to12 %>% filter(period %in% 6:12) %>% mutate(period = factor(period))

colnames(comb_df_1to12) == colnames(comb_df_tasks)

comb_df <- rbind(comb_df_1to12, comb_df_tasks)

comb_df_10 <- comb_df %>% filter(period == 10)
comb_df_8 <- comb_df %>% filter(period == 8)

table(comb_df$condition)

```



## Optional to save computational time

```{r}
entropy_summary <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/results/entropy_summary20240319.csv")
hr_rhrv_summary <- read_csv("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/results/hr_rhrv_summary_20240319.csv")

```

```{r}

idtocond_new <- pt_cond_cleaned %>% 
  mutate(condition = case_match(pi_geneticdiagnosistype,
                                0 ~ "Control",
                                1 ~ "Mutation",
                                2 ~ "Deletion")) %>% 
  mutate(id = tolower(subject_id)) %>% 
  select(id, condition)
idtocond_new <- idtocond_new[1:97,]

#write_csv(idtocond_new, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/pt_cond97.csv")

```


## Cat

```{r categories}

base_df = comb_df %>% filter(period == 1) %>% arrange(id, period, interval)
pre_df = comb_df %>% filter(period == 2) %>% arrange(id, period, interval)
task_df = comb_df %>% filter(period == 3|period == 4|period == 5) %>% arrange(id, period, interval)
post_df = comb_df %>% filter(period == 6|period == 7|period == 8|period == 9|period == 10|period == 11|period == 12) %>% arrange(id, period, interval)
all_df = comb_df %>% arrange(id, period, interval)

```

# EDA

## HR/RR/BP vs time

```{r eda}
colnames(comb_df)

comb_df <- comb_df_6_12

# HR, RR, BP vs time for all periods

ggplotly(ggplot(comb_df, aes(x = interval, y = rate_hr, color = id)) +
  geom_line() +
  theme(legend.position = "none") +
  labs(title = "Heart Rate") +
  facet_wrap(~period))

ggplotly(ggplot(comb_df, aes(x = interval, y = rate_rsp, color = id)) +
  geom_line() +
  theme(legend.position = "none") +
  labs(title = "Respiration Rate") +
  facet_wrap(~period))

ggplotly(ggplot(comb_df, aes(x = interval, y = pval_bp, color = id)) +
  geom_line() +
  theme(legend.position = "none") +
  labs(title = "Blood Pressure") +
  facet_wrap(~period))


ggplot(comb_df %>% filter(id != "mi027"), aes(x = interval, y = pval_bp, color = id)) +
  geom_line() +
  theme(legend.position = "none") +
  labs(title = "Blood Pressure over Time in Collection Periods 6-12",
       x = "Time (seconds)",
       y = "Blood Pressure (mmHg)") +
  facet_wrap(~period)




############

ggplotly(ggplot(hr_df, aes(x = pSec, y = rate, color = id)) +
  geom_line() +
  theme(legend.position = "none") +
  labs(title = "Heart Rate") +
  facet_wrap(~period))

```


# Note

Our prior knowledge of the experiment set up and the EDA shows us that the most homogenous (similar actions, same time frame) periods are 6-12, which are the "post" periods after the task. We will approach the data in the following ways:

1. Throw all of the periods in and explore what that looks like.
2. Look at metrics for periods 6-12.
3. Average the metrics of 6-12 and use that as our global metric.
  a. This is good because it seems to be the most homogeous group.
  b. This is problematic because it is during the response to a stressor task, so they aren't exactly comparable. 

# Functions for Metrics

For ease of reference and reproducibility, we will create a function for each metric, even if it's just one line. 

```{r}

# For the following, we will be using the same parameters for hr, bp, and rsp. This may change in the future. 

# Approximate and sample entropy from pracma
  # https://search.r-project.org/CRAN/refmans/pracma/html/entropy.html

metric_approx_entropy <- function(data_time_series) {
   # data_time_series should be rate_hr, rate_rsp, and pval_bp
   pracma::approx_entropy(data_time_series, edim = 2, r = 0.2*sd(data_time_series), elag = 1)
 }

metric_sample_entropy <- function(data_time_series) {
   # data_time_series should be rate_hr, rate_rsp, and pval_bp
   pracma::sample_entropy(data_time_series, edim = 2, r = 0.2*sd(data_time_series), tau = 1)
}

metric_te_old <- function(te_matrix, dir_a, dir_b){

te <- RTransferEntropy::transfer_entropy(
                te_matrix %>% ungroup %>% pull(dir_a),
                te_matrix %>% ungroup %>% pull(dir_b),
                 lx = 3,
                 ly = 3
                 )

te_ab <- te$coef[1,1]

return(te_ab)

# te_ab <- te$coef[1,1] # te a -> b
# se_ab <- te$coef[1,3] # se a -> b
# p_ab <- te$coef[1,4] # pvalue a -> b
# te_ba <- te$coef[2,1] # te b -> a
# se_ba <- te$coef[2,3] # se b -> a
# p_ba <- te$coef[2,4] # pvalue b -> a
# 
# return(c(te_ab = te_ab, 
#          se_ab = se_ab,
#          p_ab = p_ab,
#          te_ba = te_ba,
#          se_ba = se_ba,
#          p_ba = p_ba
#          ))
}

metric_te <- function(dir_a, dir_b){

te <- RTransferEntropy::transfer_entropy(
                dir_a,
                dir_b,
                 lx = 3,
                 ly = 3
                 )

te_ab <- te$coef[1,1]

return(te_ab)
}

test_te_fctn <- metric_te(comb_df_1to12 %>% filter(period == 8) %>% filter(id =="mi024"), "rate_hr", "pval_bp")

metric_corr_dim <- function(data_time_series){
#  library(nonlinearTseries)
  df_ts <- data_time_series
#old.par = par(mfrow = c(1, 2))

# tau-delay estimation based on the autocorrelation function
tau.acf = timeLag(df_ts, technique = "acf",
                  lag.max = 100, do.plot = F)
# tau-delay estimation based on the mutual information function
tau.ami = timeLag(df_ts, technique = "ami", 
                  lag.max = 100, do.plot = F)
#par(old.par)

#par(mfrow = c(1,1))

# compute embedding dimension
emb.dim = estimateEmbeddingDim(df_ts, time.lag = tau.ami,
                               max.embedding.dim = 15)

# phase space reconstruction
#tak = buildTakens(df_ts,embedding.dim = emb.dim, time.lag = tau.ami)
#scatter3D(tak[,1], tak[,2], tak[,3],
#          main = "Reconstructed Phase Space",
#          col = 1, type="o",cex = 0.3)

# Estimations

# Correlation dimension

cd = corrDim(df_ts,
             min.embedding.dim = emb.dim,
             max.embedding.dim = emb.dim + 5,
             time.lag = tau.ami, 
             min.radius = 0.001, max.radius = 50,
             n.points.radius = 40,
             do.plot=FALSE)
#plot(cd)
cd.est = estimate(cd, regression.range=c(2,7),
                  use.embeddings = 8:12)
return(cd.est)
}

```


# Approximate and sample entropy
https://search.r-project.org/CRAN/refmans/pracma/html/entropy.html

* These are for equally spaced timeseries 

```{r approx_and_sample_entropy_pracma}
library(pracma)

test <- comb_df %>% filter(period == "Post5") %>% filter(id == "mi005")
data <- pull(test, rate_hr)

comb_df <- comb_df %>% 
  mutate(rate_hr = HR) %>% 
  mutate(pval_bp = BP) %>% 
  mutate(rate_rsp = RR)

id_cond_df <- comb_df %>% ungroup() %>% select(id, condition) %>% unique()

approx_entropy(pull(test, rate_hr), edim = 2, r = 0.2*sd(pull(test, rate_hr)), elag = 1)

# DFA
dfa <- nonlinearTseries::dfa(pull(test, rate_hr), dfaOrder = 1, do.plot = FALSE)
estimate(dfa)

# MSE
# https://rdrr.io/cran/RespirAnalyzer/man/MSE.html 
library(Rlibeemd)
multiscale_entropy(data, scales = scales, method = "SampEn")


test_te_fctn <- metric_te(comb_df_6_12 %>% filter(period == 8) %>% filter(id =="mi024"), rate_hr, pval_bp)

entropy_summary_6_12 <- comb_df %>% 
  group_by(id, period) %>% 
  na.omit() %>% 
  summarize(hr_approx_entropy = metric_approx_entropy(rate_hr)#,
           # bp_approx_entropy = metric_approx_entropy(pval_bp),
          #  rsp_approx_entropy = metric_approx_entropy(rate_rsp),
          #  hr_sample_entropy = metric_sample_entropy(rate_hr),
          #  bp_sample_entropy = metric_sample_entropy(pval_bp),
          #  rsp_sample_entropy = metric_sample_entropy(rate_rsp),
          #  hr_corr_dim = metric_corr_dim(rate_hr)
            ) %>% 
  ungroup() %>% 
  left_join(id_cond_df, by = "id") %>% 
  mutate(condition = as.factor(condition)) %>% 
  select(id, period, condition, everything())

te_summary_1_12 <- comb_df_1to12 %>% 
  group_by(id, period) %>% 
  na.omit() %>% 
  summarize(te_hr_bp = metric_te(rate_hr, pval_bp) ,
            te_bp_hr = metric_te(pval_bp, rate_hr),
            te_hr_rsp = metric_te(rate_hr, rate_rsp),
            te_rsp_hr = metric_te(rate_rsp, rate_hr),
            te_bp_rsp = metric_te(pval_bp, rate_rsp),
            te_rsp_bp = metric_te(rate_rsp, pval_bp)
            ) %>% 
  ungroup() %>% 
  left_join(id_cond_df, by = "id") %>% 
  mutate(condition = as.factor(condition)) %>% 
  select(id, period, condition, everything())

write_csv(te_summary_1_12, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/te_summary_1to12_20240329.csv")


metric_summary <- comb_df %>% 
  group_by(id, period) %>% 
  na.omit() %>% 
  summarize(hr_approx_entropy = pracma::approx_entropy(rate_hr, edim = 2, r = 0.2*sd(rate_hr), elag = 1),
            bp_approx_entropy = pracma::approx_entropy(pval_bp, edim = 2, r = 0.2*sd(pval_bp), elag = 1),
            rsp_approx_entropy = pracma::approx_entropy(rate_rsp, edim = 2, r = 0.2*sd(rate_rsp), elag = 1),
            hr_sample_entropy = pracma::sample_entropy(rate_hr, edim = 2, r = 0.2*sd(rate_hr), tau = 1),
            bp_sample_entropy = pracma::sample_entropy(pval_bp, edim = 2, r = 0.2*sd(pval_bp), tau = 1),
            rsp_sample_entropy = pracma::sample_entropy(rate_rsp, edim = 2, r = 0.2*sd(rate_rsp), tau = 1),
            ) %>% 
  ungroup() %>% 
  left_join(id_cond_df, by = "id") %>% 
  mutate(condition = as.factor(condition)) %>% 
  select(id, period, condition, everything())
  
#write_csv(entropy_summary, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/results/entropy_summary20240319.csv")

entropy_summary <- metric_summary %>% 
  select(-condition) %>% 
  left_join(id_cond_df, by = "id") %>% 
  select(id, period, condition, everything())

metric_summary <- entropy_summary %>% 
  select(-condition, -period) %>% 
  left_join(hr_rhrv_summary, by = "id") %>% 
  select(id, period, condition, everything())

entropy_summary <- entropy_summary_6_12 %>% 
  filter(condition %in% c("Mutation", "Deletion", "Control"))

t.test(data = entropy_summary %>% filter(condition != "Deletion"),
       hr_approx_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Mutation"),
       hr_approx_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Deletion"),
       hr_sample_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Mutation"),
       hr_sample_entropy ~ condition) 

t.test(data = entropy_summary %>% filter(condition != "Deletion"),
       bp_approx_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Mutation"),
       bp_approx_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Deletion"),
       bp_sample_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Mutation"),
       bp_sample_entropy ~ condition) 

t.test(data = entropy_summary %>% filter(condition != "Deletion"),
       rsp_approx_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Mutation"),
       rsp_approx_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Deletion"),
       rsp_sample_entropy ~ condition)
t.test(data = entropy_summary %>% filter(condition != "Mutation"),
       rsp_sample_entropy ~ condition) 


library(lme4)
mod.RI <- lmer(bp_sample_entropy ~ condition +              # exposure
                      (1|id) +                       # random intercept for each subject
                      (1|period),
                    data = entropy_summary) 
summary(mod.RI)
summary(mod.RI)$coefficients

```


## Histograms

```{r metric_histograms}

# Get names of metrics 
metric_names <- colnames(entropy_summary %>% select(-id, -period, -condition))

library(ggplot2)
library(dplyr)

library(ggplot2)
library(dplyr)
library(tools)

histogram_plot <- function(metric){
  
  metric_sym <- rlang::sym(metric) # Convert the string metric to a symbol
  title_str <- gsub("_", " ", metric) # Make the title pretty
  title_metric <- toTitleCase(title_str)

  p <- entropy_summary %>% 
    ggplot(aes(x = !!metric_sym)) +
    labs(title = paste(title_metric, "Histogram")) +
    geom_density(alpha = 0.5)
  plot(p)
}

histogram_plot_condition <- function(metric){
  # Convert the string metric to a symbol
  metric_sym <- rlang::sym(metric)
  title_str <- gsub("_", " ", metric)
  title_metric <- toTitleCase(title_str)

  p <- entropy_summary %>% 
    ggplot(aes(x = !!metric_sym, fill = condition)) +
    labs(title = paste(title_metric, "Histogram by Condition")) +
    geom_density(alpha = 0.5)
  plot(p)
}



lapply(metric_names, histogram_plot)
lapply(metric_names, histogram_plot_condition)

for (metric in metric_names){
  histogram_plot(metric)
}


# Combine them into one figure

metric_names <- colnames(entropy_summary %>% select(-id, -period, -condition))
metric_names <- colnames(hr_rhrv_summary %>% ungroup %>% select(-id, -period, -condition))
metric_names <- colnames(te_summary %>% select(-id, -period, -condition))
metric_names <- colnames(rsp_hr_dt_summary %>% ungroup %>% select(-id, -period, -condition))

#long_summary <- hr_rhrv_summary %>% 
 # mutate(test = scale(hr_IRRR))

# long_summary <- hr_rhrv_summary %>% 
# long_summary <- entropy_summary %>% 
# long_summary <- te_summary %>% 
long_summary <- rsp_hr_dt_summary %>% 
  pivot_longer(cols = metric_names,
               names_to = "metric",
               values_to = "value")

# Histograms
long_summary %>% ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~metric, nrow = 3) +
  labs(title = "Time Variability Metrics")

# Density plots
long_summary %>% ggplot(aes(x = value)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~metric, nrow = 3) +
  labs(title = "HRV Metrics")

# Density plots by condition
long_summary %>% ggplot(aes(x = value, fill = condition)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~metric, nrow = 3) +
  labs(title = "Time Variability Metrics")

```


## Entropy Plots over Period

```{r metric_plots}

metric_plot_line <- function(metric, fig_title){
combined_summary %>% 
  ggplot(aes(x = as.numeric(period), y = {{metric}}, color = condition)) +
  geom_point() +
  labs(title = fig_title,
       x = "Time Period") +
  geom_line()
}

metric_plot_average_line <- function(metric, fig_title){
  combined_summary %>% 
    ggplot(aes(x = as.numeric(period), y = {{metric}}, color = condition)) +
    stat_summary(fun = mean, geom = "point", size = 3) +  # Plot points at the mean of each group
    stat_summary(fun = mean, geom = "line", size = 1) +   # Connect means with lines
    labs(title = fig_title,
         x = "Time Period")
}


metric_plot_average_line(metric = hr_approx_entropy, fig_title = "Heart Rate Approx Entropy (All Periods)")

metric_plot_loess <- function(metric, fig_title){
combined_summary %>% 
  ggplot(aes(x = as.numeric(period), y = {{metric}}, color = condition)) +
  geom_point() +
  labs(title = fig_title,
       x = "Time Period") +
  geom_smooth()
}

for (metric in metric_names){
  entropy_plots({{metric}})
}

entropy_plots(metric = hr_approx_entropy, fig_title = "Heart Rate Approx Entropy (All Periods)")
entropy_plots(rsp_approx_entropy, "Rsp Rate Approx Entropy (All Periods)")
entropy_plots(bp_approx_entropy, "BP Approx Entropy (All Periods)")
entropy_plots(hr_sample_entropy, "Heart Rate Sample Entropy (All Periods)")
entropy_plots(rsp_sample_entropy, "Rsp Rate Sample Entropy (All Periods)")
entropy_plots(bp_sample_entropy, "BP Sample Entropy (All Periods)")



metric_plot_loess(metric = hr_approx_entropy, fig_title = "Heart Rate Approx Entropy (All Periods)")
metric_plot_average_line(metric = hr_approx_entropy, fig_title = "Heart Rate Approx Entropy (All Periods)")

metric_variables <- colnames(combined_summary)[4:dim(combined_summary)[2]]

for (metric in metric_variables){
  plot <- metric_plot_average_line(metric = {{metric}}, fig_title = paste0(metric, " (Averaged Per Condition)"))
  print(plot)
}

metric_plot_average_line(metric = rsp_t_sd, fig_title = "Rsp t sd (All Periods)")


metric_plot_average_line_2 <- function(metric, fig_title) {
  ggplot(combined_summary, aes_string(x = "period", y = metric, group = "condition", color = "condition")) +
    geom_point(alpha = 0.4) +
    stat_summary(fun.y = mean, geom = "point", size = 3) +
    stat_summary(fun.y = mean, geom = "line", aes(group = condition), size = 1) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = fig_title, x = "Time Period")
}

metric_plot_loess_2 <- function(metric, fig_title) {
  ggplot(combined_summary, aes_string(x = "period", y = metric, group = "condition", color = "condition")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = fig_title, x = "Time Period") +
    geom_smooth()
}


for (metric in metric_variables[4]) {
  plot_line <- metric_plot_average_line_2(metric, fig_title = paste0(metric, " (Averaged Per Condition)"))
#  plot_loess <- metric_plot_loess_2(metric, fig_title = paste0(metric, " (Smoothed)"))
  
  print(plot_line)
#  print(plot_loess)
  
}

for (metric in metric_variables) {
  plot_line <- metric_plot_average_line_2(metric, fig_title = paste0(metric, " (Averaged Per Condition)"))
#  plot_loess <- metric_plot_loess_2(metric, fig_title = paste0(metric, " (Smoothed)"))

  # Make sure to print the plot if you're not directly viewing it, as ggsave saves the last plot that was displayed
  print(plot_line)
  ggsave(filename = paste0("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/results/time_plots/line_points_", metric, "_20240319.png"), plot = plot_line)

#  print(plot_loess)
#  ggsave(filename = paste0("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/results/time_plots/loess_", metric, "_20240319.png"), plot = plot_loess)
}





#######

metric_plot_loess <- function(metric, fig_title){
  # Convert string to symbol and evaluate it within the data frame context
  data_for_metric <- combined_summary %>%
    filter(!is.na(.data[[metric]]))  # Filtering NA values for the specified metric

  # Compute quantiles for dynamic y-limits
  q_low <- quantile(data_for_metric[[metric]], probs = 0.05, na.rm = TRUE)
  q_high <- quantile(data_for_metric[[metric]], probs = 0.95, na.rm = TRUE)
  
  # Add some padding
  padding <- (q_high - q_low) * 0.1
  ylims <- c(q_low - padding, q_high + padding)

  # Generate plot
  ggplot(data_for_metric, aes(x = as.numeric(period), y = .data[[metric]], color = condition)) +
    stat_summary(fun.y = mean, geom = "point", size = 3) +
    stat_summary(fun.y = mean, geom = "line", size = 1) +
    labs(title = fig_title, x = "Time Period") +
    scale_y_continuous(limits = ylims)  # Set dynamic y-axis limits
}

for (metric in metric_variables) {
  metric_plot_loess(metric, fig_title = paste0(metric, " (Averaged Per Condition)"))
}

```


# Analysis

## Summary Statistics
```{r metric_summary_stats}

# Define your custom theme
custom_theme <- theme_minimal() + 
              #  theme(text = element_text(family = "Arial", color = "#333333")) +
                scale_color_manual(values = c("#1b9e77", "#d95f02", "#7570b3"))

# Set the custom theme as the default for all ggplot objects
theme_set(custom_theme)

library(dplyr)
library(knitr)

# ICC
ICCest(id, scale(hr_rmssd), data = hr_summary_all)
ICCest(period, scale(hr_rmssd), data = hr_summary_all)
ICCest(condition, scale(hr_rmssd), data = hr_summary_all)

metric_vector <- c(hr_approx_entropy, bp_approx_entropy, rsp_approx_entropy,
                     hr_sample_entropy, bp_sample_entropy, rsp_sample_entropy)

summary_table_all <- metric_summary %>%
  group_by(period, condition) %>%
  summarise(across(c(hr_approx_entropy, bp_approx_entropy, rsp_approx_entropy,
                     hr_sample_entropy, bp_sample_entropy, rsp_sample_entropy),
                   list(mean = mean, 
                        sd = sd,
                        iqr = IQR), na.rm = TRUE)) %>%
  ungroup()

# View the summary table
kable(summary_table)

metric_names <- c("hr_approx_entropy", "bp_approx_entropy", "rsp_approx_entropy",
                     "hr_sample_entropy", "bp_sample_entropy", "rsp_sample_entropy")

rsp_hr_dt_summary %>% pivot_longer(cols = metric_names,
                                names_to = "metric",
                                values_to = "value") %>% 
  ggplot(aes(x = metric, y = value, color = condition)) +
  geom_violin() +
  labs(title = "Time Variability Violin Plot") +
 # scale_c
  theme(axis.text.x = element_text(angle = 45))
  
##### Correlation
  
  library(reshape2)

correlation_matrix <-  cor(metric_summary %>% select(metric_names))
  
melted_correlation_matrix <- melt(correlation_matrix)
ggplot(melted_correlation_matrix, aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(fill = "Correlation")

ggplot(melted_correlation_matrix, aes(Var1, Var2, fill = value)) +
    geom_tile() +  # Create the tiles for heatmap
    geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 3) +  # Add text labels
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +  # Set the color gradient
    theme_minimal() +  # Use a minimal theme
    theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
          axis.title = element_blank()) +  # Remove axis titles
    labs(fill = "Correlation")  # Label the legend


```

## EDA Visualizations -- Heat Map

```{r}

library(dplyr)
library(knitr)

# Calculate mean and standard deviation for each measure across conditions and periods
summary_table <- combined_summary %>%
  group_by(period, condition) %>%
  summarise(across(c(hr_approx_entropy, bp_approx_entropy, rsp_approx_entropy,
                     hr_sample_entropy, bp_sample_entropy, rsp_sample_entropy),
                   list(mean = mean, sd = sd), na.rm = TRUE)) %>%
  ungroup()

# View the summary table
knitr::kable(summary_table)
library(ggplot2)
library(reshape2)


# Prepare the data for heatmap: we'll use hr_approx_entropy as an example
heatmap_data <- combined_summary %>%
  select(id, period, condition, hr_approx_entropy) %>%
  group_by(period, condition) %>%
  summarise(mean_hr_approx_entropy = mean(hr_approx_entropy, na.rm = TRUE)) %>%
  ungroup()

# Reshape the data for the heatmap
heatmap_matrix <- dcast(heatmap_data, period ~ condition, value.var = "mean_hr_approx_entropy")

# Create the heatmap
ggplot(heatmap_data, aes(x = condition, y = period, fill = mean_hr_approx_entropy)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = paste0("Heatmap of HR Approx Entropy"), x = "Condition", y = "Period") +
  theme_minimal()

# bp_approx_entropy
# Prepare the data for heatmap
heatmap_data <- combined_summary %>%
  select(id, period, condition, bp_approx_entropy) %>%
  group_by(period, condition) %>%
  summarise(mean_bp_approx_entropy = mean(bp_approx_entropy, na.rm = TRUE)) %>%
  ungroup()

# Reshape the data for the heatmap
heatmap_matrix <- dcast(heatmap_data, period ~ condition, value.var = "mean_bp_approx_entropy")

# Create the heatmap
ggplot(heatmap_data, aes(x = condition, y = period, fill = mean_bp_approx_entropy)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = paste0("Heatmap of BP Approx Entropy"), x = "Condition", y = "Period") +
  theme_minimal()

# rsp_approx_entropy
# Prepare the data for heatmap
heatmap_data <- combined_summary %>%
  select(id, period, condition, rsp_approx_entropy) %>%
  group_by(period, condition) %>%
  summarise(mean_rsp_approx_entropy = mean(rsp_approx_entropy, na.rm = TRUE)) %>%
  ungroup()

# Reshape the data for the heatmap
heatmap_matrix <- dcast(heatmap_data, period ~ condition, value.var = "mean_rsp_approx_entropy")

# Create the heatmap
ggplot(heatmap_data, aes(x = condition, y = period, fill = mean_rsp_approx_entropy)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = paste0("Heatmap of Resp Approx Entropy"), x = "Condition", y = "Period") +
  theme_minimal()



# function

metric_heatmap <- function(metric){
  
 # metric <-  {{metric}}
 # fig_title <- paste0(metric, " (Averaged Per Condition)")
  
  # Prepare the data for heatmap
heatmap_data <- combined_summary %>%
 # filter(!is.na(.data[[metric]])) %>% 
  select(id, period, condition, .data[[metric]]) %>%
  group_by(period, condition) %>%
  summarise(mean_metric = mean(.data[[metric]], na.rm = TRUE)) %>%
  ungroup()

# Reshape the data for the heatmap
#heatmap_matrix <- dcast(heatmap_data, period ~ condition, value.var = "mean_metric")

# Create the heatmap
ggplot(heatmap_data, aes_string(x = "condition", y = "period", fill = "mean_metric")) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = paste0("Heatmap of ", metric), x = "Condition", y = "Period") +
  theme_minimal()
}


metric_heatmap("bp_approx_entropy")

library(dplyr)
library(ggplot2)
library(reshape2)  # For dcast

metric_heatmap <- function(metric){
  
  # Assuming 'combined_summary' is a dataframe available in your environment
  # Prepare the data for heatmap
  heatmap_data <- combined_summary %>%
    filter(!is.na(all_of(metric))) %>%  # Updated to use all_of() for safe column selection
    select(id, period, condition, all_of(metric)) %>%
    group_by(period, condition) %>%
    summarise(mean_metric = mean(.data[[metric]], na.rm = TRUE), .groups = 'drop')

  # Reshape the data for the heatmap
  heatmap_matrix <- dcast(heatmap_data, period ~ condition, value.var = "mean_metric")

  # Check if 'condition' and 'period' are in the columns of heatmap_matrix
  # if (!("condition" %in% names(heatmap_matrix)) || !("period" %in% names(heatmap_matrix))) {
  #   stop("The reshaped data does not contain 'condition' or 'period' columns. Please check your data.")
  # }

  # Create the heatmap
  ggplot(heatmap_data, aes_string(x = "condition", y = "period", fill = "mean_metric")) +
    geom_tile() +
    scale_fill_gradient(low = "blue", high = "red") +
    labs(title = paste0("Heatmap of ", metric), x = "Condition", y = "Period") +
    theme_minimal()

}

# Example usage:
# metric_heatmap("bp_approx_entropy")

for (metric in metric_variables) {
  plot_heatmap <- metric_heatmap(metric)

  # Make sure to print the plot if you're not directly viewing it, as ggsave saves the last plot that was displayed
  print(plot_heatmap)
#  ggsave(filename = paste0("/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/results/time_plots/heatmap_", metric, "_20240319.png"), plot = plot_line)
}

```


# Heart and Respiration Variability Metrics

## rmssd and sd

```{r hr_rsp_rmssd_sd, eval = FALSE}

# Variability

# RMSSD --- root mean square of succesive differences
rmssd <- function(data_delta_t){
  sq <- data_delta_t^2
  avgofsq <- mean(sq)
  rmssd <- sqrt(avgofsq)
  return(rmssd)
}

# PNN50 --- Percentage of successive RR intervals that differ by more than 50 ms
pnn50 <- function(data_delta_t){
  under_50ms = 0
  over_50ms = 0
  for (i in 1:(length(data_delta_t) - 1)) {
    if(abs(data_delta_t[i] - data_delta_t[i+1]) >= 0.05){
      over_50ms = over_50ms + 1
    } else {
      under_50ms = under_50ms + 1
    }
  }
  percent = over_50ms / (over_50ms + under_50ms)
  return(percent)
}
#pnn50(dt_test)

sdrr <- function(data_delta_t){
  sd(data_delta_t)
}

rsp_df <- rsp_raw #this might be bad

hr_df_dt <- hr_df %>% 
  ungroup() %>% 
  mutate(delta_t_hr = pSec - lag(pSec)) %>% 
  mutate(delta_t_hr = if_else(abs(delta_t_hr) > 3, NA, delta_t_hr)) 
rsp_df_dt <- rsp_df %>% 
  ungroup() %>% 
  mutate(delta_t_rsp_p = pSec - lag(pSec),
         delta_t_rsp_t = tSec - lag(tSec)) %>% 
  mutate(delta_t_rsp_p = if_else(abs(delta_t_rsp_p) > 20, NA, delta_t_rsp_p)) %>% 
  mutate(delta_t_rsp_t = if_else(abs(delta_t_rsp_t) > 20, NA, delta_t_rsp_t)) 

hr_summary <- hr_df_dt %>% 
  select(-"condition") %>% 
  group_by(id, period) %>% 
  na.omit() %>% 
  summarize(#hr_dt_mean = mean(delta_t_hr),
            #hr_dt_sd = sd(delta_t_hr),
            hr_rmssd = rmssd(delta_t_hr),
         #   hr_pnn50 = pnn50(delta_t_hr),
            hr_sdrr = sd(delta_t_hr)) %>% 
  ungroup() %>% 
  left_join(id_cond_df, by = "id") %>% 
  mutate(condition = as.factor(condition))

rsp_summary <- rsp_df_dt %>% 
  select(-"condition") %>% 
  group_by(id, period) %>% 
  na.omit() %>% 
  summarize(rsp_p_rmssd = rmssd(delta_t_rsp_p),
            rsp_t_rmssd = rmssd(delta_t_rsp_t),
            #pnn50 analog?
            rsp_p_sd = sd(delta_t_rsp_p, na.rm = TRUE),
            rsp_t_sd = sd(delta_t_rsp_t, na.rm = TRUE)) %>% 
  ungroup() %>% 
  left_join(id_cond_df, by = "id") %>% 
  mutate(condition = as.factor(condition))

rsp_hr_dt_summary <- full_join(hr_summary, rsp_summary, by = c("id", "condition", "period")) %>% 
  select("id", "period", "condition", everything()) %>% 
  filter(id != "mi031") 

#write_csv(rsp_hr_dt_summary, "/Users/christinekuryla/Documents/A_Columbia_Research/MiSBIE/misbie/data/results/rsp_hr_dt_summary_20240417.csv")


rsp_hr_dt_summary_612 <- rsp_hr_dt_summary %>% 
  filter(period %in% 6:12)

hist(hr_summary$hr_dt_sd)
hist(hr_summary$hr_dt_mean)
hist(hr_summary$hr_rmssd)
hist(hr_summary$hr_pnn50)


# t-test for hr delta t sd
t.test(data = hr_summary %>% filter(condition == "Control" | condition == "Mutation"), hr_dt_sd ~ condition)
t.test(data = hr_summary %>% filter(condition == "Control" | condition == "Deletion"), hr_dt_sd ~ condition)
t.test(data = hr_summary %>% filter(condition == "Mutation" | condition == "Deletion"), hr_dt_sd ~ condition)

# t-test for hr delta t mean
t.test(data = hr_summary %>% filter(condition == "Control" | condition == "Mutation"), hr_dt_mean ~ condition)
t.test(data = hr_summary %>% filter(condition == "Control" | condition == "Deletion"), hr_dt_mean ~ condition)
t.test(data = hr_summary %>% filter(condition == "Mutation" | condition == "Deletion"), hr_dt_mean ~ condition)

# t-test for hr rmssd
t.test(data = hr_summary %>% filter(condition == "Control" | condition == "Mutation"), hr_rmssd ~ condition)
t.test(data = hr_summary %>% filter(condition == "Control" | condition == "Deletion"), hr_rmssd ~ condition)
t.test(data = hr_summary %>% filter(condition == "Mutation" | condition == "Deletion"), hr_rmssd ~ condition)

# t-test for hr pnn50
t.test(data = hr_summary %>% filter(condition == "Control" | condition == "Mutation") %>% filter(period %in% 1:12), hr_pnn50 ~ condition)
t.test(data = hr_summary %>% filter(condition == "Control" | condition == "Deletion"), hr_pnn50 ~ condition)
t.test(data = hr_summary %>% filter(condition == "Mutation" | condition == "Deletion"), hr_pnn50 ~ condition)

#hr_summary <- hr_summary_all %>% filter(period == 2)

# Wilcoxon rank sum test
wilcox.test(data = hr_summary %>% filter(condition == "Control" | condition == "Mutation"), hr_pnn50 ~ condition)
wilcox.test(data = hr_summary %>% filter(condition == "Control" | condition == "Deletion"), hr_pnn50 ~ condition)
wilcox.test(data = hr_summary %>% filter(condition == "Deletion" | condition == "Mutation"), hr_pnn50 ~ condition)

wilcox.test(data = hr_summary %>% filter(condition == "Control" | condition == "Mutation") %>% filter(period == "8"),
            hr_pnn50 ~ condition)

```


```{r}

pca_prep <- hr_summary %>% 
  pivot_wider(names_from = period,
              values_from = hr_dt_sd, # hr_rmssd, hr_dt_mean, hr_dt_sd
              id_cols = id) %>% 
  filter(!(id %in% c("mi014","mi031","mi042"))) %>% 
  column_to_rownames(var = "id") %>% 
  select(-C3, -Acquisition) %>% 
  na.omit() 

# pca_prep <- hr_summary %>% 
#   pivot_wider(names_from = id,
#               values_from = hr_dt_sd, # hr_rmssd, hr_dt_mean, hr_dt_sd
#               id_cols = period) %>% 
#  # filter(!(id %in% c("mi014","mi031","mi042"))) %>% 
#   column_to_rownames(var = "period") %>% 
# #  select(-C3, -Acquisition) %>% 
#   na.omit() 

pca_result <- prcomp(pca_prep, center = TRUE, scale = TRUE)
view(pca_result)
loadings(pca_result)

id_to_group <- hr_summary %>% 
  select(id, condition) %>% 
  unique()
pca_id_to_group <- id_to_group %>% 
  filter(id %in% row.names(pca_prep))

# loading library 
library(ggfortify) 
pca_plot <- autoplot(pca_result, 
						data = pca_id_to_group, 
						colour = 'condition') 

pca_plot

```



## https://cran.r-project.org/web/packages/fractalRegression/fractalRegression.pdf 

# Sample Entropy

* A comprehensive comparison and overview of R packages for calculating sample entropy (Chen 2019)
  * https://academic.oup.com/biomethods/article/4/1/bpz016/5634143 


# Multiscale Entropy (MSE)

```{r}


```


```{r}
hr_summary_all <- hr_summary_all %>% filter(!(period %in% c("Acquisition", "C3", "5")))

# ICC
ICCest(id, scale(hr_rmssd), data = hr_summary_all)
ICCest(period, scale(hr_rmssd), data = hr_summary_all)
ICCest(condition, scale(hr_rmssd), data = hr_summary_all)

ICCest(id, scale(hr_dt_sd), data = hr_summary_all)
ICCest(period, scale(hr_dt_sd), data = hr_summary_all)
ICCest(condition, scale(hr_dt_sd), data = hr_summary_all)

ICCest(id, scale(hr_dt_mean), data = hr_summary_all)
ICCest(period, scale(hr_dt_mean), data = hr_summary_all)
ICCest(condition, scale(hr_dt_mean), data = hr_summary_all)

ICCest(id, scale(hr_pnn50), data = hr_summary_all)
ICCest(period, scale(hr_pnn50), data = hr_summary_all)
ICCest(condition, scale(hr_pnn50), data = hr_summary_all)
```


```{r}
# To easily look at output
coeff_md_vs_control <- function(model){
  
beta.RI.deletion <- summary(model)$coefficients[2,1] #[(c+1),1]
se.RI.deletion   <- summary(model)$coefficients[2,2] #[(c+1),2]
lci.RI.deletion  <- beta.RI.deletion - 1.96*se.RI.deletion
uci.RI.deletion  <- beta.RI.deletion + 1.96*se.RI.deletion
coef.RI.deletion <- paste0(round(beta.RI.deletion,3), " (95% CI: ", round(lci.RI.deletion,3), 
                     ", ", round(uci.RI.deletion,3), ")")
coef.RI.deletion 

beta.RI.mutation <- summary(model)$coefficients[3,1] #[(c+1),1]
se.RI.mutation   <- summary(model)$coefficients[3,2] #[(c+1),2]
lci.RI.mutation  <- beta.RI.mutation - 1.96*se.RI.mutation
uci.RI.mutation  <- beta.RI.mutation + 1.96*se.RI.mutation
coef.RI.mutation <- paste0(round(beta.RI.mutation,3), " (95% CI: ", round(lci.RI.mutation,3), 
                     ", ", round(uci.RI.mutation,3), ")")
coef.RI.mutation 
  
output_string <- paste0("Deletion vs Control: ", coef.RI.deletion, 
                        "\n", "Mutation vs Control: ", coef.RI.mutation)

cat(output_string)

}
```


```{r}
mod.RI <- lmer(hr_pnn50 ~ condition +              # exposure
                 #     period +
                 (1|period) + #
                      (1|id),                       # random intercept for each subject
                    data = hr_summary_all) 
summary(mod.RI)
summary(mod.RI)$coefficients


beta.RI.pm <- summary(mod.RI)$coefficients[3,1]
se.RI.pm   <- summary(mod.RI)$coefficients[3,2]
lci.RI.pm  <- beta.RI.pm - 1.96*se.RI.pm
uci.RI.pm  <- beta.RI.pm + 1.96*se.RI.pm

coef.RI.pm <- paste0(round(beta.RI.pm,3), " (95% CI: ", round(lci.RI.pm,3), 
                     ", ", round(uci.RI.pm,3), ")")
coef.RI.pm 



```



```{r}

# This is what happens when we don't use mixed models
summary(lm(scale(hr_pnn50) ~ condition, data = hr_summary_all))
summary(lm(scale(hr_rmssd) ~ condition + period, data = hr_summary_all))
summary(lm(scale(hr_dt_sd) ~ condition, data = hr_summary_all))
summary(lm(scale(hr_dt_mean) ~ condition, data = hr_summary_all))

mod.RI <- lmer(scale(hr_pnn50) ~ condition +   
                 #     period +
                 (1|period) + #
                      (1|id),     # random intercept for each subject
                    data = hr_summary_all) 
summary(mod.RI)
summary(mod.RI)$coefficients
coeff_md_vs_control(mod.RI)


mod_ri_pnn50 <- lmer(scale(hr_pnn50) ~ condition +   
                 (1|period) + (1|id),     
                 data = hr_summary_all) 
summary(mod_ri_pnn50)
summary(mod_ri_pnn50)$coefficients
coeff_md_vs_control(mod_ri_pnn50)

mod_ri_rmssd <- lmer(scale(hr_rmssd) ~ condition +   
                 (1|period) + (1|id),     
                 data = hr_summary_all) 
summary(mod_ri_rmssd)
summary(mod_ri_rmssd)$coefficients
coeff_md_vs_control(mod_ri_rmssd)

mod_ri_sd <- lmer(scale(hr_dt_sd) ~ condition +   
                 (1|period) + (1|id),     
                 data = hr_summary_all) 
coeff_md_vs_control(mod_ri_sd)

mod_ri_pnn50 <- lmer(scale(hr_pnn50) ~ condition +   
                 (1|period) + (1|id),     
                 data = hr_summary_all) 
coeff_md_vs_control(mod_ri_pnn50)

```


```{r post}

# Post is period 6-12 so I want to try to combine them

hr_df_post <- hr_df %>% 
  filter(period %in% c("6", "7", "8", "9", "10", "11", "12")) %>% 
  mutate(period = as.numeric(period)) %>% 
  arrange(id, period, pSec)

hr_summary_post <- hr_df_post %>% 
  group_by(id) %>% 
  na.omit() %>% 
  summarize(hr_dt_mean = mean(delta_t),
            hr_dt_sd = sd(delta_t),
            hr_rmssd = rmssd(delta_t),
            hr_pnn50 = pnn50(delta_t)) %>% 
  ungroup() %>% 
  left_join(id_cond_df, by = "id") %>% 
  mutate(condition = as.factor(condition))

```

```{r}

hr_summary_all %>% 
  ggplot(aes(x = hr_pnn50)) + 
  geom_boxplot(aes(color = condition), alpha = 0.2, outlier.shape = NA) +
  #geom_jitter(aes(color = condition), size=0.4, alpha=0.6) +
  facet_grid(~ period, scales = "free") #+
  
 # theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +             
#  theme(strip.text.x = element_text(size = 7), strip.text.y = element_text(size = 7))  # Change font size


```

```{r}

library(frac)

```

