---
title: "Transfer Entropy"
author: "Christine Lucille Kuryla"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load and wrangle data

```{r load_data}

comb_df <- read_csv(file.path("./data_processed", "comb_df_20240524.csv"))

```


```{r}

library(tidyverse)

# somehow load comb_df

# string vector
periods_str <- c("Base",
                 "Pre",
                 "Instruct",
                 "PrepTask",
                 "Task",
                 "Post5",
                 "Post10",
                 "Post20",
                 "Post30",
                 "Post60",
                 "Post90",
                 "Post120",
                 "DBT",
                 "ST",
                 "SST",
                 "CP")

# factor vector
# create a vector with the periods as factors in the correct order (not alphabetical)
# this is for ease of plotting later
periods_factor <- factor(periods_str, levels = periods_str)

ids <- comb_df %>% ungroup() %>% select(id) %>% unique()

comb_df <- comb_df %>% 
  mutate(period = factor(period, levels = periods_str))

comb_df_apen <- comb_df
comb_df <- comb_df_apen

```

In order to calculate transfer entropy of the differences, we need the differences in the time values. Let's append them to the combined dataframe.

```{r}

#comb_df_loaded <- comb_df
comb_df <- comb_df_loaded

# comb_df <- comb_df %>%
#   group_by(id, period) %>%
#   mutate(interval = factor(interval, levels = unique(interval))) %>%
#   mutate(interval = as.integer(interval))

comb_df_8 <- comb_df_8 %>% 
  group_by(id, period) %>%
  mutate(rate_diff = HR - lag(HR),
         pval_diff = BP - lag(BP),
         pval_rsp_diff = rsp_pVal - lag(rsp_pVal),
         rate_rsp_diff = RR - lag(RR)) %>%
  rename(rate_hr = HR, rate_hr_diff = rate_diff,
         pval_bp = BP, pval_bp_diff = pval_diff,
         pval_rsp = rsp_pVal, rate_rsp = RR) %>%
  mutate(interval = as.integer(interval)) %>% 
  drop_na()


comb_df_pre_te <- comb_df

```

# Change period definitions if desired

> periods
 [1] "Base"     "Pre"      "Instruct" "PrepTask" "Task"     "Post5"    "Post10"   "Post20"  
 [9] "Post30"   "Post60"   "Post90"   "Post120"  "DBT"      "ST"       "SST"      "CP"      
> ### Base (period 1)
> ### pre (period 2)
> ### Task (period 3-5)
> ### post (period 6-12)

```{r consolidate_periods}

library(dplyr)
library(forcats)

# Helper function to extract numeric part of period
extract_numeric <- function(x) {
  as.numeric(gsub("[^0-9]", "", x))
}

# Update the code with custom ordering and creation of interval_longer
comb_df_8 <- comb_df %>%
  mutate(period = factor(period, levels = c("Base", "Pre", "Instruct", "PrepTask", "Task", 
                  "Post5", "Post10", "Post20", "Post30", "Post60", "Post90", "Post120", 
                  "DBT", "ST", "SST", "CP" ))) %>%
  arrange(id, period) %>% 
  mutate(period_longer = case_match(period,
                                    "Base" ~ "Base",
                                    "Pre" ~ "Pre",
                                    c("Instruct", "PrepTask", "Task") ~ "Task",
                                    c("Post5", "Post10", "Post20", "Post30", "Post60", "Post90", "Post120") ~ "Post",
                                    .default = as.character(period))) %>%
  mutate(interval = as.numeric(interval)) %>%  # Ensure interval is numeric
  arrange(id, period_longer,
          case_when(
            period %in% c("Instruct", "PrepTask", "Task") ~ as.numeric(factor(period, levels = c("Instruct", "PrepTask", "Task"))),
            TRUE ~ extract_numeric(as.character(period))
          ),
          interval) %>%
  group_by(id, period_longer) %>%
  mutate(interval_longer = row_number()) %>%
  #mutate(interval_longer = as.integer(interval)) %>%
  ungroup() %>% 
  mutate(period_longer = factor(period_longer, levels = c("Base", "Pre", "Task", 
                  "Post", "DBT", "ST", "SST", "CP")))

# Display the resulting data frame
head(comb_df_8)

ggplot(comb_df_8 %>% filter(period_longer == "Post", id == "mi018"), aes(x = interval_longer, y = HR, group = id, color = period)) +
  geom_line() +
#  facet_wrap(~ period_longer) + 
  ggtitle("Heart Rate")

ggplot(comb_df_8, aes(x = interval_longer, y = RR, group = id, color = condition)) +
  geom_line() +
  facet_wrap(~ period_longer) + ggtitle("Respiration Rate")

ggplot(comb_df, aes(x = interval, y = BP, group = id, color = condition)) +
  geom_line() +
  facet_wrap(~ period) + ggtitle("Blood Pressure")



View(comb_df_8 %>% filter(id == "mi005", period_longer == "Post") %>% select(id, period, interval, period_longer, interval_longer))

comb_df_8_orig <- comb_df_8
comb_df_8 <- comb_df_8 %>% 
  group_by(id, period) %>%
  mutate(rate_diff = HR - lag(HR),
         pval_diff = BP - lag(BP),
         pval_rsp_diff = rsp_pVal - lag(rsp_pVal),
         rate_rsp_diff = RR - lag(RR)) %>%
  rename(rate_hr = HR, 
    rate_hr_diff = rate_diff,
         pval_bp = BP, 
    pval_bp_diff = pval_diff,
         pval_rsp = rsp_pVal, rate_rsp = RR) %>%
  mutate(interval_orig = interval) %>% 
  mutate(interval = as.integer(interval)) 

```



# Calculate TE

```{r te_for_1}


# set comb_df, periods_str

library(RTransferEntropy)
library(future)

#cur <- comb_df %>% filter(id == "mi069", period == "Post")

future::plan(multisession)

# Initialize data frames to store results
te_df <- data.frame()
skipped_te <- data.frame()

id_old <- c("mi001", "mi002", "mi005", "mi006", "mi007", "mi008", "mi009", "mi010", "mi011", "mi012", "mi013", "mi015" ,"mi016","mi017","mi018", "mi019", "mi020", "mi021", "mi022", "mi023", "mi024", "mi025", "mi026","mi027", "mi028", "mi029","mi030", "mi036", "mi037", "mi040" ,"mi041", "mi043" ,"mi044", "mi045", "mi046", "mi048", "mi049", "mi050", "mi051","mi052" ,"mi053", "mi054", "mi055" ,"mi056", "mi057" ,"mi059", "mi060")

# Loop through ids and periods (this takes a while!)

#for(id_str in ids %>% pull(id)) {
for(id_str in id_old) {
# for(period_str in periods_str){
#for(period_str in unique(comb_df_8$period_longer)){

for(period_str in c("Post", "SST", "ST")){

# 
# for(id_str in c("mi021", "mi053", "mi069", "mi100", "mi053", "mi037")) {
#        for(period_str in c("Post60", "PrepTask", "Post5")){
         
  #  cur = comb_df %>% filter(id == id_str, period == period_str)
    cur = comb_df_8 %>% filter(id == id_str, period_longer == period_str) %>% mutate(period = period_longer)
    
    print(id_str)
    print(period_str)
    
    # Skip if empty
    if (dim(cur)[1] == 0) {
      skipped_instance <- data.frame(skipped_id = id_str,
                                     skipped_period = period_str,
                                     reason = "null")
      skipped_te <- rbind(skipped_te, skipped_instance)
      next
    }
    
    # Calculate TE within tryCatch
    tryCatch({
      
      # TE on differences
      TE1hb_diff = transfer_entropy(cur$rate_hr_diff, cur$pval_bp_diff, type = 'quantiles', lx = 1, ly = 1)#, nboot = 0)#, burn = 20)
      TE2hr_diff = transfer_entropy(cur$rate_hr_diff, cur$rate_rsp_diff, type = 'quantiles', lx = 1, ly = 1)#, nboot = 0)#, burn = 20)
      TE3br_diff = transfer_entropy(cur$pval_bp_diff, cur$rate_rsp_diff, type = 'quantiles', lx = 1, ly = 1)#, nboot = 0)#, burn = 20)
      
      # TE on original time series
      TE1hb = transfer_entropy(cur$rate_hr, cur$pval_bp, type = 'quantiles', lx = 1, ly = 1)#, burn = 20)
      TE2hr = transfer_entropy(cur$rate_hr, cur$rate_rsp, type = 'quantiles', lx = 1, ly = 1)#, burn = 20)
      TE3br = transfer_entropy(cur$pval_bp, cur$rate_rsp, type = 'quantiles', lx = 1, ly = 1)#, burn = 20)
      
      te_row <- data.frame(
        id = id_str,
        period = period_str,
      # condition = condition,  
        te_hr_bp_diff = TE1hb_diff$coef[1, 1],
        te_bp_hr_diff = TE1hb_diff$coef[2, 1],
        te_hr_rr_diff = TE2hr_diff$coef[1, 1],
        te_rr_hr_diff = TE2hr_diff$coef[2, 1],
        te_bp_rr_diff = TE3br_diff$coef[1, 1],
        te_rr_bp_diff = TE3br_diff$coef[2, 1],
        te_hr_bp = TE1hb$coef[1, 1],
        te_bp_hr = TE1hb$coef[2, 1],
        te_hr_rr = TE2hr$coef[1, 1],
        te_rr_hr = TE2hr$coef[2, 1],
        te_bp_rr = TE3br$coef[1, 1],
        te_rr_bp = TE3br$coef[2, 1]#,
        # te_hr_bp_diff_p = TE1hb_diff$coef[1, 4],
        # te_bp_hr_diff_p = TE1hb_diff$coef[2, 4],
        # te_hr_rr_diff_p = TE2hr_diff$coef[1, 4],
        # te_rr_hr_diff_p = TE2hr_diff$coef[2, 4],
        # te_bp_rr_diff_p = TE3br_diff$coef[1, 4],
        # te_rr_bp_diff_p = TE3br_diff$coef[2, 4]#,
        # te_hr_bp_p = TE1hb$coef[1, 4],
        # te_bp_hr_p = TE1hb$coef[2, 4],
        # te_hr_rr_p = TE2hr$coef[1, 4],
        # te_rr_hr_p = TE2hr$coef[2, 4],
        # te_bp_rr_p = TE3br$coef[1, 4],
        # te_rr_bp_p = TE3br$coef[2, 4]
      )
      
      te_df <- rbind(te_df, te_row)
      
    }, error = function(e) {
      # Handle the error by storing the skipped instance
      skipped_instance <- data.frame(skipped_id = id_str,
                                     skipped_period = period_str,
                                     reason = "error")
      skipped_te <- rbind(skipped_te, skipped_instance)
      
      # Print error message for debugging
      print(paste("Error for id:", id_str, "and period:", period_str))
      print(e)
    })
  }
}


print(te_df)
print(skipped_te)

#colnames(te_df) <- names(te_row)

# add condition back in 
id_cond <- comb_df %>% 
  ungroup() %>% 
  select(id, condition) %>% 
  unique()

te_df <- left_join(te_df, id_cond, by = "id") %>% 
  select(id, period, condition, everything())

plan(sequential)
# 
# write_csv(te_df, file = file.path("./data_processed", "te_df_notdetrended_20240524.csv"))
# write_csv(skipped_te, file = file.path("./data_processed", "skipped_te_notdetrended_20240524.csv"))

te_df_8 <- te_df
#write_csv(te_df, file = file.path("./data_processed", "te_df_8_notdetrended_20240524.csv"))
#write_csv(skipped_te, file = file.path("./data_processed", "skipped_te_8_notdetrended_20240524.csv"))

te_df_l_1 <- te_df_8
te_df_l_3 <- te_df

```

Many were skipped, let's explore that. 

```{r te_completeness}

# Checking completeness
te_df %>% 
  group_by(id) %>% 
  summarize(n_periods = n()) %>% 
  group_by(n_periods) %>% 
  summarize(n = n())

te_df %>% 
  group_by(period) %>% 
  summarize(n = n())

#   n_periods     n
#       <int> <int>
# 1         5     1
# 2         9     1
# 3        10     1
# 4        11     3
# 5        12     4
# 6        13     5
# 7        14    25
# 8        15    50
# 9        16    15

# Checking completeness
te_df %>% 
  group_by(id) %>% 
  filter(period == "PrepTask") %>% 
  summarize(n_periods = n()) %>% 
#  filter(n_periods == 15) %>% 
  summarize(n = n())

# 34 ----> many,  many PrepTasks failed

te_df %>% 
  group_by(id, period) %>% 
  summarize(n = n())
  
  
te_df_0 <- read_csv(file.path("./data_processed", "te_df_notdetrended_20240524.csv"))


```

# Comparing TE values (first pass)

# T tests and Wilcoxon Signed Rank tests (Mann-Whitney)

```{r}

#head(all)
library(tidyr)
library(broom)

# Assuming df has columns: id, condition, period, and metric

combined_summary <- te_df

all_metric_names <- colnames(combined_summary %>% select("te_hr_bp_diff":"te_rr_bp_diff"))
#all_metric_names <- colnames(combined_summary %>% select("te_hr_bp_diff":"te_rr_bp"))

# all_metric_names <- colnames(combined_summary %>% 
#                                select(-"id", -"condition", -"period"))

# Assuming df has columns: id, condition, period, and multiple metrics
# all_metric_names is your vector of metric names
library(dplyr)
library(tidyr)

results_list <- lapply(all_metric_names, function(metric) {
  metric_data <- combined_summary %>%
    select(id, condition, period, !!sym(metric)) %>%
    rename(metric_value = !!sym(metric))
  
  pairwise_comparisons <- metric_data %>%
    group_by(period) %>%
    do({
      subset_data <- .
      comparisons <- expand.grid(unique(subset_data$condition), unique(subset_data$condition))
      comparisons <- comparisons[comparisons$Var1 != comparisons$Var2,]
      
      # Initialize results dataframe
      results <- data.frame()
      
      for (i in 1:nrow(comparisons)) {
        cond1 <- comparisons$Var1[i]
        cond2 <- comparisons$Var2[i]
        
        data1 <- subset(subset_data, condition == cond1)$metric_value
        data2 <- subset(subset_data, condition == cond2)$metric_value
        
        if (length(data1) > 1 && length(data2) > 1) {
          t_test_result <- t.test(data1, data2)
          wilcoxon_test_result <- wilcox.test(data1, data2, exact = FALSE)  # allow approximate p-value
          
          results <- rbind(results, data.frame(
            Var1 = cond1,
            Var2 = cond2,
            period = unique(subset_data$period),
            mean_1 = t_test_result$estimate[1],
            mean_2 = t_test_result$estimate[2],
            n_1 = length(data1),
            n_2 = length(data2),
            t_test_p = t_test_result$p.value,
            wilcoxon_test_p = wilcoxon_test_result$p.value,
            sign_g1_v_g2 = if_else(sign(t_test_result$estimate[1] - t_test_result$estimate[2]) == 1, paste0(cond1, " higher than ", cond2), paste0(cond1, " lower than ", cond2))
          ))
        }
      }
      results
    }) %>%
    unnest(cols = everything())  # Specify columns to unnest
  
  group_means <- metric_data %>%
    group_by(condition, period) %>%
    summarise(mean_metric = mean(metric_value, na.rm = TRUE), .groups = 'drop')
  
  # Combine results
  results <- left_join(pairwise_comparisons, group_means, by = c("Var1" = "condition", "period"))
  results$metric <- metric  # Add the metric name to the results
  
  return(results)
})

results_list <- bind_rows(results_list)  # Combine all results into one data frame

results_list <- results_list %>% 
  filter(Var1 != "MELAS" & Var2 != "MELAS") %>% 
  filter(!(Var1 == "Control" & Var2 == "Mutation")) %>% 
  filter(!(Var1 == "Control" & Var2 == "Deletion")) %>% 
  filter(!(Var1 == "Deletion" & Var2 == "Mutation")) %>% 
  rename(Group1 = Var1) %>% 
  rename(Group2 = Var2) %>% 
  select(Group1, Group2, period, metric, sign_g1_v_g2, t_test_p, wilcoxon_test_p, everything())

p_val_per_period <- results_list

results_t_sig <- results_list %>% 
  filter(t_test_p <= 0.15)

results_wilcoxon_sig <- results_list %>% 
  filter(wilcoxon_test_p <= 0.15)

# write_csv(p_val_per_period, file = file.path("./data_processed", "p_val_per_period_8_notdetrended_108_20240603.csv"))

```

