---
title: "Oct.15 starting trying"
output: html_document
date: "2022-10-16"
---
```{r}
library(tidyverse)
library(pracma)
library(ggplot2)
library(egg)
library(gridExtra)
```

## heart rate

```{r}
load("hr_cleaned.RData")
hr_df = hr_cleaned
```


#### check the first stage - base
```{r}
hr_1 = hr_df %>% filter(period == 1) %>%
  group_by(id) %>%
  mutate(diff = rate - lag(rate)) 

ggplot(hr_1, aes(x = time, y = rate, group = id)) + 
  geom_smooth(se = FALSE, size = 0.2) # rate plot
ggplot(hr_1, aes(x = time, y = diff, group = id)) + 
  geom_smooth(se = FALSE, size = 0.2) # diff plot

## kernel density of heart rate and the diff
par(mfrow=c(1,2))
a <- rnorm(200, mean = 72, sd = 12)   
plot(density(hr_1$rate, na.rm = T))
lines(density(a), col = "red")  

b <- rnorm(200, mean = 0, sd = 2.2)   
plot(density(hr_1$diff, na.rm = T))
lines(density(b), col = "red") # heavy tail??

```


#### check the fifth stage - pre-task
```{r}
hr_5 = hr_df %>% filter(period == 5) %>%
  group_by(id) %>%
  mutate(diff = rate - lag(rate)) 

ggplot(hr_5, aes(x = time, y = rate, group = id)) + 
  geom_smooth(se = FALSE, size = 0.2) # rate plot
ggplot(hr_5, aes(x = time, y = diff, group = id)) + 
  geom_smooth(se = FALSE, size = 0.2) # diff plot

## kernel density of heart rate and the diff
par(mfrow=c(1,2))
a <- rnorm(100000, mean = 72, sd = 12)   
plot(density(hr_1$rate, na.rm = T))
lines(density(a), col = "red") 

b <- rnorm(100000, mean = 0, sd = 2.2)   
plot(density(hr_1$diff, na.rm = T))
lines(density(b), col = "red") # heavy tail?
```


### trying functions from time series

```{r}
mi020 = hr_df %>% filter(id == "mi020")
mi020_hr = ts(as.data.frame(mi020$rate))
mi023 = hr_df %>% filter(id == "mi023")
mi023_hr = ts(as.data.frame(mi023$rate))

a = autoplot(mi023_hr)
b = autoplot(mi020_hr)
grid.arrange(a,b, nrow = 2)

```




## respiratory

```{r}
load("rsp_cleaned.RData")
rsp_df = rsp_cleaned
```


#### check the first stage - base
```{r}
rsp_1 = rsp_df %>% filter(period == 1) %>%
  group_by(id) %>%
  mutate(diff = rate - lag(rate)) %>%
  filter(id != "mi057") 
# filter out mi057 for abnormal much more measurements than others

ggplot(rsp_1, aes(x = time, y = rate, group = id)) + 
  geom_smooth(se = FALSE, size = 0.2) # rate plot

ggplot(rsp_1, aes(x = time, y = diff, group = id)) + 
  geom_smooth(se = FALSE, size = 0.2) # diff plot

## kernel density of heart rate and the diff
par(mfrow=c(1,2))
a <- rnorm(200000, mean = 17.5, sd = 3.5)   
plot(density(rsp_1$rate, na.rm = T))
lines(density(a), col = "red")  

b <- rnorm(200000, mean = 0, sd = 2)   
plot(density(rsp_1$diff, na.rm = T))
lines(density(b), col = "red") # heavy tail??

```


#### check the fifth stage - pre-task
```{r}
rsp_5 = rsp_df %>% filter(period == 5) %>%
  group_by(id) %>%
  mutate(diff = rate - lag(rate)) %>%
  filter(id != "mi057") 
# filter out mi057 for abnormal much more measurements than others

ggplot(rsp_5, aes(x = time, y = rate, group = id)) + 
  geom_smooth(se = FALSE, size = 0.2) # rate plot

ggplot(rsp_5, aes(x = time, y = diff, group = id)) + 
  geom_smooth(se = FALSE, size = 0.2) # diff plot

## kernel density of heart rate and the diff
par(mfrow=c(1,2))
a <- rnorm(200000, mean = 17.5, sd = 3.5)   
plot(density(rsp_1$rate, na.rm = T))
lines(density(a), col = "red")  

b <- rnorm(200000, mean = 0, sd = 2)   
plot(density(rsp_1$diff, na.rm = T))
lines(density(b), col = "red") # heavy tail??

```


#### trying functions for time serise

```{r}
mi020 = rsp_df %>% filter(id == "mi020")
mi020_rsp = ts(as.data.frame(mi020$rate))
mi023 = rsp_df %>% filter(id == "mi023")
mi023_rsp = ts(as.data.frame(mi023$rate))

a = autoplot(mi023_rsp)
b = autoplot(mi020_rsp)
grid.arrange(a,b, nrow = 2)


```




## problems
* no patient condition for mi045 to mi072

* time window: every 5 seconds per value.

* paramertric likelihood assumption (condition probability)

* auto-regression model (m-lags, num of lag sufficient) - ar1, ar2, ar3, ar4 





```{r}
-2.5 + 1.96 * 1.15
```

