---
title: "Sensitivity Analysis Master Document"
author: "Christine Lucille Kuryla"
date: "2024-07-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*** think about the TE results that have a large p value
*** TE vs effective TE


vignette: https://cran.r-project.org/web/packages/RTransferEntropy/vignettes/transfer-entropy.html 
paper: https://www.sciencedirect.com/science/article/pii/S2352711019300779 
reference manual: https://cran.r-project.org/web/packages/RTransferEntropy/RTransferEntropy.pdf 

# Goals and Summary

This project involves varying multiple parts of the TE pipeline, including:
* Data processing
  1. Order of data cleaning
  2. Time interval of interpolation
* Transfer Entropy calculations
  1. Varying parameters
  2. Differences vs non-differences

We will do this by using functions and varying their order and parameters for ease of interpretation.

We will then compare the results of the TE analysis for each potential combination.

# Order of data cleaning

Rmd: ""

## Components



## Orders






```{r load_libraries}

library(tidyverse)
library(RTransferEntropy)

```

```{r load_data_and_process}

```

